### 2018-06-16 LSFA/first_steps_into_forcing.tex
template for LSFA paper

\documentclass[9pt]{entcs} \usepackage{entcsmacro}
\usepackage{graphicx}
\sloppy

\input{header-draft}
\def\lastname{Gunther, Pagano, S\'anchez Terraf}
\begin{document}
\begin{frontmatter}
  \title{First steps towards a formalization of Forcing}
  \author{Emmanuel Gunther%\thanksref{ALL}
    \thanksref{myemail}}
  \address{FaMAF\\ Universidad Nacional de C\'ordoba\\
    C\'ordoba, Argentina} \author{Miguel Pagano\thanksref{coemail}}
  \address{FaMAF\\Universidad Nacional de C\'ordoba\\
    C\'ordoba, Argentina}
  \author{Pedro S\'anchez Terraf\thanksref{co2email}}
  \address{CIEM-FaMAF\\Universidad Nacional de C\'ordoba\\
    C\'ordoba, Argentina}
 \thanks[myemail]{Email:
    \href{mailto:gunther@famaf.unc.edu.ar} {\texttt{\normalshape
        gunther@famaf.unc.edu.ar}}} \thanks[coemail]{Email:
    \href{mailto:pagano@famaf.unc.edu.ar} {\texttt{\normalshape
        pagano@famaf.unc.edu.ar}}}  \thanks[co2email]{Email:
    \href{mailto:sterraf@famaf.unc.edu.ar} {\texttt{\normalshape
        sterraf@famaf.unc.edu.ar}}} 
 \thanks[ALL]{Supported by Secyt-UNC project 33620180100465CB.} 
\begin{abstract} 
\end{abstract}
\begin{keyword}
Isabelle/ZF, forcing, preorder, Rasiowa-Sikorski lemma, names, generic extension.
\end{keyword}
\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\subsection{Models of ZFC}

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\subsection{Isabelle/ZF}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Forcing posets}\label{sec:forcing-posets}

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\subsection{The Rasiowa-Sikorski lemma}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The generic extension}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Names}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work}
\begin{itemize}
\item Develop an interface between Paulson's relativization results
  and countable models $M$ of $\ZFC$. This will show
  that every such $M$ is closed under well-founded recursion and, in
  particular, that contains names for each of its
  elements. Consequently, $M\sbq M[G]$.
\item Prove that the basic axioms of $\ZFC$ hold in the generic
  extension. 
\item After prototyping the forcing relation, prove the Axiom Scheme
  of Separation (the first that needs to use the machinery of forcing
  nontrivially).
\end{itemize}

\end{document}