### 2018-06-22 LSFA/intro.tex
forcing

\section{Introduction}\label{sec:introduction}
Set Theory play a double role in Mathematics: It is one of its
possible foundations, and also an active research area.
%
As it is widely known, Georg Cantor introduced its main concepts and in
particular showed the fundamental result that the real line, $\R$  is not
coordinable with the natural numbers. Soon after this, he posed the
most important question in the field, written as a conjecture:
\begin{quote}
  The \emph{Continuum Hypothesis} ($\CH$). Every uncountable subset of $\R$ is
  coordinable with $\R$.
\end{quote}

The current  axiomatic foundation of Set Theory is through 
first-order logic and uses the axioms by  Zermelo, Fraenkel, and
including the Axiom of Choice ($\AC$) among them. This theory is
known by the $\ZFC$ acronym. G\"odel showed that $\CH$ can't be
refuted using the $\ZFC$, unless the latter are inconsistent (we say
that \emph{$\CH$ is relatively 
  consistent with  $\ZFC$}). For a while, this result left the
possibility that one might be able show that $\ZFC\models \CH$, but in
a groundbreaking work \cite{Cohen-CH-PNAS}, Paul Cohen discovered the
technique of 
\emph{forcing} and showed that $\neg\CH$ is relatively consistent with
$\ZFC$. Forcing has since then been used for showing innumerable
independence results and to perform mathematical constructions.

\subsection{Models of $\ZFC$}
By G\"odel's Second Incompleteness Theorem, we can't show prove that
there exists a model of $\ZFC$. More formally, if we assume that
mathematical proofs can be encoded as theorems of $\ZFC$ and that
the latter do not lead to contradictions (i.e., $\ZFC$ is
\emph{consistent}), then we can't prove that there exists a set $M$
and a binary relation $E$ such that $\lb M,E\rb$ satisfies the $\ZFC$
axioms.

A relative consistency proof for an axiom $A$ is then obtained by assuming that such a
$\lb M,E\rb$ exists, and constructing another $\lb M',E'\rb$
satisfying $\ZFC + A$. We single out a very special kind of models:
%
\begin{definition}\label{def:transitive-model}
  \begin{enumerate}
  \item A set $M$ (of sets) is \emph{transitive} if for all $x\in M$ and 
    $y\in x$, we have $y\in M$ (i.e., every element of $M$ is a subset
    of $M$).
  \item $\lb M,E\rb$ is a \emph{transitive model} if $M$ is transitive
    set and  $E$ is the membership relation $\in$ restricted to
    $M$. It is \emph{countable} if $M$ is coordinable with a subset of
    $\N$; we then say that $M$ is a \emph{ctm}. 
  \end{enumerate}
\end{definition}
%
As in the last sentence, one usually uses the underlying set to refer to a transitive
model, since the relation is fixed.

Despite G\"odel's Second, we can find transitive models for every
finite fragment of $\ZFC$. More precisely, 
%
\begin{theorem}\label{th:ctm-finite-axioms}
  For each finite subset
  $\Phi\sbq \ZFC$, the statement \emph{``there exists a countable
    transitive model of $\Phi$''} is a theorem of $\ZFC$.
\end{theorem}
%
This follows by a
combination of three results (the Reflection Principle,
L\"owenheim-Skolem Theorem, and the Mostowksi Collapse). The reader
can consult the details in \cite{kunen2011set}. Consistency arguments that assume
the existence of a ctm $M$ of $\ZFC$ can usually be replaced by a
model as in Theorem~\ref{th:ctm-finite-axioms}, since a first-order
proof (e.g.\ of contradiction) involves only finitely many axioms.

Assume that $M$ is a ctm of $\ZFC$. In a portrayal convenient for this
introduction, G\"odel showed that $M$ contains a minimal submodel
$L^M$ of the same ``height'' (i.e.\ having the same ordinals)
that satisfies $\ZFC+\CH$. The sets in $L^M$ are called
\emph{constructible} and are in a sense ``definable.''

It is a primary need to have a means to correlate  (first-order)
properties satisfied by a model $M$ and those of a 
submodel $N\sbq M$. As a simple example on this, consider 
%% $M\defi \{\emptyset,\{\emptyset\},\{\{\emptyset\}\}\}$ and
%% $N\defi\{\{\emptyset\},\{\{\emptyset\}\}\}$, and let 
%% $\phi(x)\defi \neg\exists y (y \in x)$. It is clear that 
%% $M\models \phi(\emptyset)$; but we have
%% \[
%% M\not\models \phi(\{\emptyset\}) \quad\text{ but }\quad N\models \phi(\{\emptyset\}).
%% \]
%% There is a discrepancy between what $M$ and $N$ \emph{believe} about
%% their common element $\{\emptyset\}$.
$M\defi \{a,b,c, \{a,b\},\{a,b,c\}\}$ and
$N\defi\{a,b,\{a,b,c\}\}$, and let 
\[
\phi(x,y,z)\defi \forall w \,( w\in z \lsii w=x \o w=y).
\]
%% It is clear that  $M\models \phi(\{a,b\})$; but we have
Then we have
\[
M\not\models \phi(a,b,\{a,b,c\}) \quad\text{ but }\quad N\models \phi(a,b,\{a,b,c\}).
\]
There is a discrepancy between  $M$ and $N$ about $\{a,b,c\}$ being ``the
(unordered) pair of $a$ and $b$.'' We say that $\phi$ holds for
$a,b,\{a,b,c\}$ \emph{relative} to $N$. It is immediate to see that
$\phi$ holds  for $x,y,z$ relative to $N$ if and only if
\[
\phi^N(x,y,z)\defi \forall w.\ w\in N\limp ( w\in z \lsii w=x \o w=y)
\] 
holds. $\phi^N$ is called the \emph{relativization of $\phi$ to
  $N$}. One can generalize this operation of relativization to the
class of all sets satisfying a first-order predicate $C$ in a
straightforward way:
\[
\phi^C(x,y,z)\defi \forall w.\ C(w)\limp ( w\in z \lsii w=x \o w=y)
\]

It can be shown elementarily that if $M$ and $N$ are transitive,
$\phi^N$ holds if and only if $\phi^M$ holds,  for $x,y,z\in N$. We
say then that $\phi$ is \emph{absolute between $N$ and $M$.}

The concepts of relativization and absoluteness are central to task of
transferring truth of axioms in $M$ to  $L^M$.

\subsection{Forcing}
Forcing is a technique to extend countable transitive models of
$\ZFC$. This process is guaranteed to preserve the $\ZFC$
axioms but allows to fine-tune what other first order properties the
extension will have. Given a ctm $M$ of $\ZFC$, a new set $G$ is
adjoined to $M$ and under some hypothesis ($G$ being ``generic''),
there exists a smallest model $M[G]$ of $\ZFC$ that includes $M$ and
contains $G$.

\fbox{The next should be spread accross the other sections}

The easiest way to define genericity is by using preorders.
Given a preorder with top $\lb\PP,\leq,\1\rb$ in $M$, define
\emph{filter} $G$ ... , \emph{dense} ...  $G$ is an $M$-\emph{generic}
if it intersects every dense subset of $\PP$ that lies in $M$.

But since $M$ is countable,

The  Rasiowa-Sikorski theorem  states that that for every countable 
family of dense subsets there is a filter that intersects them
all. Thus, there are generic filters $G$ for countable transitive models.  In
general, no such  $G$ belongs to $M$ and therefore the extension
$M[G]$ is proper.

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% TeX-master: "first_steps_into_forcing"
%%% End: