### 2018-06-24 LSFA/isabelle-zf.tex
compatibilizo relativizacion con la intro

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%
\section{Isabelle/ZF}

Let us introduce briefly Paulson's formalization of ZF
\cite{paulson2017isabelle} in Isabelle and aspects of his formal proof
for the relative consistency of the Axiom of Choice
\cite{paulson_2003}; we will only focus on those aspects that are
essential to keep this paper self-contained, and refer the interested
reader to Paulson's articles. 
%% Isabelle/ZF is a theory built upon the
%% core library FOL of classical first order logic.
Isabelle/ZF includes a development of classical first order logic,
FOL. Both of them are  built upon the core library \emph{Pure}.

In Isabelle/ZF sets are \emph{individuals}, that is terms of type
\isatt{i} and formulas have type \isatt{o} 
(akin to a \emph{Bool} type, but at the object level). 
As in usual presentations
of ZF, Isabelle/ZF has only one relation symbol
\isatt{\ensuremath{\in}} (equality and inclusion are defined
using membership); it also features one constant \isatt{\isadigit{0}}
for the empty set and \isatt{inf} for an infinite set and some other
set constructors, again terms of type \isatt{i}, to ease the
developments. The axioms, being formulas, are terms of type \isatt{o};
the foundation axiom, for example, is formalized as:
\isa{{\isachardoublequoteopen}A\ {\isacharequal}\ {\isadigit{0}}\
  {\isasymor}%
  \ {\isacharparenleft}{\isasymexists}x{\isasymin}A{\isachardot}\ %
  {\isasymforall}y{\isasymin}x{\isachardot}\ %
  y{\isasymnotin}A{\isacharparenright}%
  {\isachardoublequoteclose}}. Besides the axioms, Isabelle/ZF also
introduces several definitions (for example, pairs and sets defined by
comprehension using separation) and syntactic abbreviations to keep
the formalization close to the customary manner of doing mathematics.
Working with the library and extendig it is quite
straightforward. 
As an example, we introduce  a new term-former which is combination of
instances of replacement and separation to speak of the image of a function over some set
defined by comprehension, namely $\{b(x): x\in A\text{ and }Q(x)\}$:
%% Since we find a need to combine the replacement and
%% separation schemes to speak of the image of a function over some set
%% defined by comprehension, we introduced a new term-former:
\begin{isabelle}
  \isacommand{definition}\isamarkupfalse%
  \ SepReplace\ {\isacharcolon}{\isacharcolon}\
  {\isachardoublequoteopen}{\isacharbrackleft}i{\isacharcomma}\
  i{\isasymRightarrow}i{\isacharcomma}\ i{\isasymRightarrow}\
  o{\isacharbrackright}\
  {\isasymRightarrow}i{\isachardoublequoteclose}\
  \isakeyword{where}\isanewline \ \
  {\isachardoublequoteopen}SepReplace{\isacharparenleft}A{\isacharcomma}b{\isacharcomma}Q{\isacharparenright}\
  {\isacharequal}{\isacharequal}\ {\isacharbraceleft}y\ {\isachardot}\
  x{\isasymin}A{\isacharcomma}\
  y{\isacharequal}b{\isacharparenleft}x{\isacharparenright}\
  {\isasymand}\
  Q{\isacharparenleft}x{\isacharparenright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}
\noindent %% It is more convenient to add an abbreviation for a simpler
%% writing and a nicer reading, so we use
We are then able to add an abbreviation resembling the usual
mathematical writing by setting 
\isa{{\isacharbraceleft}b\
  {\isachardot}{\isachardot}\ x{\isasymin}A{\isacharcomma}\
  Q{\isacharbraceright}} as a notation for
\isa{SepReplace{\isacharparenleft}A{\isacharcomma}b{\isacharcomma}Q{\isacharparenright}}. The
characterization of our new constructor is given by
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ Sep{\isacharunderscore}and{\isacharunderscore}Replace{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbraceleft}b{\isacharparenleft}x{\isacharparenright}\ {\isachardot}{\isachardot}\ x{\isasymin}A{\isacharcomma}\ P{\isacharparenleft}x{\isacharparenright}\ {\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}b{\isacharparenleft}x{\isacharparenright}\ {\isachardot}\ x{\isasymin}{\isacharbraceleft}y{\isasymin}A{\isachardot}\ P{\isacharparenleft}y{\isacharparenright}{\isacharbraceright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

%% We are not discussing how, if possible at all, to express that some
%% set satisfies some axiom. For example, we might well be interested in
%% proving $\emptyset \models \mathit{foundation}$, but there is no
%% sensible way to state this, without encoding formulas and axioms in
%% ZFC.\footnote{Indeed Paulson has encoded formulas using a very general
%%   method to encode recursive definitions. In our development, although
%%   not reported in this paper, we use this internalization.}
%% 
%% \fbox{ You need this in order to formalize $\models$ as a predicate of
%%   type $i\ent i \ent o$. (Don't know how to put that here.)
%% }

%% In
%% Isabelle/ZF there are, however, relativized versions of the
%% axioms. This alternative statement of the axioms are relativized with
%% respect to classes, which in Isabelle/ZF corresponds to predicates over
%% sets
We now discuss relativization in Isabelle/ZF. Relativized versions of the
 axioms can be found in the formalization of constructibility. For
 example, the relativized Axiom of Foundation is
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\ foundation{\isacharunderscore}ax\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}i{\isacharequal}{\isachargreater}o{\isacharparenright}\ {\isacharequal}{\isachargreater}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}foundation{\isacharunderscore}ax{\isacharparenleft}M{\isacharparenright}\ {\isacharequal}{\isacharequal}\isanewline
\ \ \ \ \ \ \ \ {\isasymforall}x{\isacharbrackleft}M{\isacharbrackright}{\isachardot}\ {\isacharparenleft}{\isasymexists}y{\isacharbrackleft}M{\isacharbrackright}{\isachardot}\ y{\isasymin}x{\isacharparenright}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymexists}y{\isacharbrackleft}M{\isacharbrackright}{\isachardot}\ y{\isasymin}x\ {\isacharampersand}\ {\isachartilde}{\isacharparenleft}{\isasymexists}z{\isacharbrackleft}M{\isacharbrackright}{\isachardot}\ z{\isasymin}x\ {\isacharampersand}\ z\ {\isasymin}\ y{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

% ∀x[c]
\noindent The relativized quantifier
\isa{{\isasymforall}x{\isacharbrackleft}M{\isacharbrackright}{\isachardot}\
  P(x)} is a shorthand for \isa{{\isasymforall}x{\isachardot}\ M(x)
  {\isasymlongrightarrow} P(x)}. In order to express that a (set)
satisfies this axiom we use  the coercion
\isa{{\isacharhash}{\isacharhash} :: i ={\isachargreater} (i
  ={\isachargreater} o)} provided by Isabelle/ZF. As a trivial example we
can show that the empty set satisfies Foundation:
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ emp{\isacharunderscore}foundation\ {\isacharcolon}\ {\isachardoublequoteopen}foundation{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}{\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

% Let us ponder if one can state, let alone prove, that some set \isa{M
%   :: i} satisfies a statement \isa{\ensuremath{\phi} :: o}. Since
% \isa{o} does not have an inductive definition, we cannot define the
% satisfaction of \isa{\ensuremath{\phi}}.

% As we have outlined above, relativization is a key concept when
% proving that axioms satisfied by some model $M$ are also satisfied in
% another model constructed from $M$. In this paper, we show the complete
% proof that if a set \isa{M :: i} satisfies the

Mathematical texts usually start by fixing the context which defines
parameters and assumptions necessary to develop theorems
and results. In Isabelle the way of defining contexts is through
\emph{locales}~\cite{ballarin2010tutorial}.
Locales can be combined and extended adding more parameters and assuming
more facts, leading to a new locale. For example a context describing
lattices can be extended to distributive lattices.
The way to instantiate a locale is \emph{interpreting} it, which consists
of giving concrete values to parameters and proving the assumptions.
In our work, we use locales to organize the formalization and to make
explicit the assumptions of the most important results.

% locales

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% TeX-master: "first_steps_into_forcing"
%%% End: