### 2018-06-25 LSFA/names.tex
Modificaciones menores en names.tex

\subsection{Names}
\label{sec:names}
In this section we formalize the function $\val$ that allows to
construct the elements of the generic extension $M[G]$ from elements
of the ctm $M$ and the generic filter $G$. The definition of $\val$
can be written succintly as
%
\begin{equation}\label{eq:def-val}
\val(G,\tau)\defi \{\val(G,\sigma) : \exists p\in \PP
(\lb\sigma,p\rb\in\tau \y p\in G)\}.
\end{equation}
%
As with any recursive definition, to justify that the definition makes
sense (i.e., it defines anything at all) one has to explicit the
well-founded relation which is used. In this case, we use the relation
$\mathit{ed}$:
\[
x \mathrel{\mathit{ed}} y \iff \exists p . \lb x,p\rb\in y.
\]
Recall that in $\ZFC$, an ordered pair $\lb x,y \rb$ is the set
$\{\{x\},\{x,y\}\}$. It is trivial to deduce the well-foundedness of
$\mathit{ed}$ from the fact that $\in$ is well-founded, which follows
from the Foundation Axiom.

The  formalization of $\val$ in Isabelle/ZF involves the function
\isa{wfrec}, which takes as an argument a well-founded relation of
type \isa{i}, i.e.\ a set of pairs. Therefore we restrict
$\mathit{ed}$ above to a set.
%
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ edrel\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}edrel{\isacharparenleft}A{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}{\isacharless}x{\isacharcomma}y{\isachargreater}\ {\isasymin}\ A{\isacharasterisk}A\ {\isachardot}\ x\ {\isasymin}\ domain{\isacharparenleft}y{\isacharparenright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}
%
Since \isa{edrel(A)} is a subset of a  well-founded relation (the
transitive closure of the membership relation restricted to \isa{A}),
then it is well-founded as well.

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ edrel{\isacharunderscore}sub{\isacharunderscore}memrel{\isacharcolon}\ {\isachardoublequoteopen}edrel{\isacharparenleft}A{\isacharparenright}\ {\isasymsubseteq}\ trancl{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
\dots
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}edrel\ {\isacharcolon}\ {\isachardoublequoteopen}wf{\isacharparenleft}edrel{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ wf{\isacharunderscore}subset\ {\isacharbrackleft}of\ {\isachardoublequoteopen}trancl{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}edrel{\isacharunderscore}sub{\isacharunderscore}memrel\ wf{\isacharunderscore}trancl\ wf{\isacharunderscore}Memrel{\isacharparenright}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\end{isabelle}

\noindent We now turn to the definition of $\val$. In order to use
\isa{wfrec} we need to define a functional \isa{Hv}.

\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Hv\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hv{\isacharparenleft}G{\isacharcomma}x{\isacharcomma}f{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}\ f{\isacharbackquote}y\ {\isachardot}{\isachardot}\ y{\isasymin}\ domain{\isacharparenleft}x{\isacharparenright}{\isacharcomma}\ {\isasymexists}p{\isasymin}P{\isachardot}\ {\isacharless}y{\isacharcomma}p{\isachargreater}\ {\isasymin}\ x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ val\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}\ {\isacharequal}{\isacharequal}\ wfrec{\isacharparenleft}edrel{\isacharparenleft}eclose{\isacharparenleft}M{\isacharparenright}{\isacharparenright}{\isacharcomma}\ {\isasymtau}{\isacharcomma}\ Hv{\isacharparenleft}G{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
Then we can recover the recursive expression~(\ref{eq:def-val}) thanks to the
following lemma:
%
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}val{\isacharcolon}
\isanewline
\ \ \ \ \ \ \ \  {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ val{\isacharparenleft}G{\isacharcomma}x{\isacharparenright}\ {\isacharequal}\ {\isacharbraceleft}val{\isacharparenleft}G{\isacharcomma}t{\isacharparenright}\ {\isachardot}{\isachardot}\ t{\isasymin}domain{\isacharparenleft}x{\isacharparenright}\ {\isacharcomma}\ {\isasymexists}p{\isasymin}P\ {\isachardot}\ \ {\isasymlangle}t{\isacharcomma}\ p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

We can finally define the generic extension of $M$ by $G$, also
setting up the notation $M[G]$ for it:
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ GenExt\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \ \ \ \ {\isacharparenleft}{\isachardoublequoteopen}M{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}\ {\isadigit{9}}{\isadigit{0}}{\isacharparenright}\ \isakeyword{where} \isanewline
\ \ {\isachardoublequoteopen}GenExt{\isacharparenleft}G{\isacharparenright}{\isacharequal}{\isacharequal}\ {\isacharbraceleft}val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}{\isachardot}\ {\isasymtau}\ {\isasymin}\ M{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

%% It is to be noted that this results holds under the assumption that
%% the name \isa{t} is in $M$.

We now provide names for elements in $M$. That is, for each $x\in M$,
we define $\chk(x)$ (usually denoted by $\check{x}$ in the literature)
such that $\val(G,\chk(x))=x$. This will show that $M\sbq M[G]$, with
a caveat we make explicit in the end of this section. As explained in
the introduction, the fact that $M[G]$ extends $M$ is crucial to show
that $\ZFC$ holds in the former.
%
The definition of $\chk(x)$ is a straightforward $\in$-recursion:
\begin{equation}
  \label{eq:def-check}
  \chk(x)\defi\{\lb\chk(y),\1\rb : y\in x\}
\end{equation}
Now the set-relation argument for \isa{wfrec} is the membership
relation restricted to a set \isa{A}, \isa{Memrel(A)}.

\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ Hcheck\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}i{\isacharcomma}i{\isacharbrackright}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hcheck{\isacharparenleft}z{\isacharcomma}f{\isacharparenright}\ \ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}\ {\isacharless}f{\isacharbackquote}y{\isacharcomma}one{\isachargreater}\ {\isachardot}\ y\ {\isasymin}\ z{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ check\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}check{\isacharparenleft}x{\isacharparenright}\ {\isacharequal}{\isacharequal}\ wfrec{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}{\isacharbraceleft}x{\isacharbraceright}{\isacharparenright}{\isacharparenright}{\isacharcomma}\ x\ {\isacharcomma}\ Hcheck{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
Here, \isa{eclose} returns the (downward) $\in$-closure of its
argument. The main result is stated as follows:

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ valcheck\ {\isacharcolon}\ {\isachardoublequoteopen}y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ {\isasymforall}x{\isasymin}M{\isachardot}\ check{\isacharparenleft}x{\isacharparenright}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ val{\isacharparenleft}G{\isacharcomma}check{\isacharparenleft}y{\isacharparenright}{\isacharparenright}\ \ {\isacharequal}\ y{\isachardoublequoteclose}
\end{isabelle}

On the one hand, the only requirement on the set \isa{G} is that it
contains the top element of the poset \isa{P}. On the other hand, it
is necessary that all the relevant names are indeed in $M$ (i.e.,
$\chk(x) \in M$ if $x\in M$). It requires a serious development to
fulfill this assumption. One of the hardest parts of Paulson's
formalization of constructibility involves showing that models are
closed under recursive construction. We will eventually formalize that
if $M\models\ZFC$ and the arguments of \isa{wfrec} are in $M$, then
its value also is. This will require to adapt to ctm models several
locales defined in \cite{paulson_2003} that were intended to be used
for the class of constructible sets.

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% TeX-master: "first_steps_into_forcing"
%%% End: