### 2018-07-13 LSFA/arXiv/first_steps_into_forcing.tex
typo

\documentclass{article}
\usepackage{isabelle,isabellesym}
\usepackage{hyperref}
\usepackage[numbers]{natbib}

\input{header-draft}

\begin{document}
\title{First steps towards a formalization of Forcing}
\author{Emmanuel Gunther
  \and 
  Miguel Pagano
  \and 
  Pedro S\'anchez Terraf}
\maketitle

\begin{abstract} 
  We lay the ground for an Isabelle/ZF formalization of Cohen's technique of
  \emph{forcing}. We formalize the definition of forcing notions as
  preorders with top, dense subsets, and generic filters. We formalize
  a version of the principle of Dependent Choices and using it
  we prove the Rasiowa-Sikorski lemma on the existence of generic filters.

  Given a transitive set $M$, we define its generic extension $M[G]$,
  the canonical names for elements of $M$, and finally show that if $M$
  satisfies the axiom of pairing, then $M[G]$ also does.
\end{abstract}
\keywords{
Isabelle/ZF, forcing, preorder, Rasiowa-Sikorski lemma, names, generic extension.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{intro}

\input{isabelle-zf}

\input{posets}

\input{rs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{gen-ext}

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%--%-%-%-%-%-%-%-%
\input{names}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{pair-gen-ext}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{conclusion}

\bibliographystyle{mi-estilo-else}
\bibliography{draft}
%\input{first_steps_into_forcing.bbl}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: