### 2018-08-18 LSFA/related.tex
final sin revisar, falta Jaber

\subsection{Related work}\label{sec:related-work}
Before we proceed, we will discuss very succinctly recent related
formalizations.

There is another major framework in Isabelle based on higher order
logic, Isabelle/HOL. This framework is very active, and as a 
consequence more automated tools are available. Isabelle/HOL has 
basic chapters on set theory. One of those, by Steven Obua, proceeds up to
well founded relations and provides translations between types in HOL
(for instance \isatt{nat}) to  sets  (elements of type
\isatt{ZF}). Another one, by A.~Popescu and D.~Traytel, reaches
cardinal artihmetic. This is fairly limited for our purposes.

Concerning automation, B.~Zhan has developed a new tool called
\texttt{auto2} and applied it to untyped set theory
\cite{10.1007/978-3-319-66107-0_32}. He has redeveloped essentially
the basic results in Isabelle/ZF, but goes in a different
direction. Nevertheless, a majority of results in
Isabelle/ZF are not yet implemented using this tool, and a another
downside is that proofs using it do not follow the standard Isar
language.

There are some forcing-related formalizations in the theory
of types. Most relevant is the work by K.~Quirin \cite{Quirin}, where
a sheaf-theoretic initial approach to forcing is implemented in
Coq. This language is extremely different to the standard approach of
constructing models of $\ZFC$, and it might be difficult (once the
forcing machinery is set) to translate results in the literature using
ctms to this one. In any case, the translation to set theory of what
Quirin accomplishes is to define a generic extension (where $\CH$
should fail) and to construct a set $K$ (a candidate counterexample)
and injections $\N\hookrightarrow K$ and $K\hookrightarrow\R$. But the
most important part, that is, that there are no surjections 
$\N\twoheadrightarrow K$ and $K\twoheadrightarrow\R$, is left for a future
work. % Finally,  Jaber et al.\ \cite{jaber:hal-01319066}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% TeX-master: "first_steps_into_forcing"
%%% End: