### 2018-11-07 LSFA/ENTCS/pair-gen-ext.tex

\section{Pairing in the generic extension}
\label{sec:pairing-generic-extension}
Given a predicate
\isa{C\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}},
we can consider the \emph{class} of all sets \isa{x} such that
\isa{C(x)}. In \cite{paulson_2003} a large corpus of set-theoretical
concepts are expressed in relative form.

In the case of the Pairing Axiom the following definition in 
Paulson's formalization captures
what a class \isa{C} believes about a set being the unordered pair of
other two:
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ upair\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}i{\isasymRightarrow}o{\isacharcomma}i{\isacharcomma}i{\isacharcomma}i{\isacharbrackright}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}upair{\isacharparenleft}C{\isacharcomma}a{\isacharcomma}b{\isacharcomma}z{\isacharparenright}\ {\isacharequal}{\isacharequal}\ a\ {\isasymin}\ z\ {\isasymand}\ b\ {\isasymin}\ z\ {\isasymand}\ {\isacharparenleft}{\isasymforall}x{\isacharbrackleft}C{\isacharbrackright}{\isachardot}\ x{\isasymin}z\ {\isasymlongrightarrow}\ x\ {\isacharequal}\ a\ {\isasymor}\ x\ {\isacharequal}\ b{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
%
Then the fact that Pairing axiom holds in \isa{C} is stated as follows.
%
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ upair{\isacharunderscore}ax\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}i{\isasymRightarrow}o{\isacharparenright}\ {\isasymRightarrow}\ o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ \ \ {\isachardoublequoteopen}upair{\isacharunderscore}ax{\isacharparenleft}C{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isasymforall}x{\isacharbrackleft}C{\isacharbrackright}{\isachardot}\ {\isasymforall}y{\isacharbrackleft}C{\isacharbrackright}{\isachardot}\ {\isasymexists}z{\isacharbrackleft}C{\isacharbrackright}{\isachardot}\ upair{\isacharparenleft}C{\isacharcomma}x{\isacharcomma}y{\isacharcomma}z{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

In order use the \isa{upair{\isacharunderscore}ax} for a set model, we
use the \isatt{\#\#}  operator.  We state the main result of this
section in the context \isatt{forcing{\isacharunderscore}data}.  
%
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ pair{\isacharunderscore}preserv\ {\isacharcolon}\ \isanewline
\ \ {\isachardoublequoteopen}one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ upair{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharparenright}\ {\isasymLongrightarrow}\ upair{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharbrackleft}G{\isacharbrackright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

Let $x$ and $y$ be elements in $M[G]$. By definition of the generic extension, there exist
elements $\tau$ and $\rho$ in $M$ such that $x = \val(G,\tau)$ and
$y = \val(G,\rho)$.  We need to find an element in $M[G]$ that contains exactly
these elements; for that we should construct a name $\sigma\in M$ such that
$\val(G,\sigma) = \{ \val(G,\tau) , \val(G,\rho) \}$.

The candidate, motivated by the definition of $\chk$,  is
$\sigma = \{\langle \tau , \mathrm{one} \rangle , \langle \rho , \mathrm{one} \rangle \}$. 
Our remaining tasks are to show 
\begin{enumerate}
  \item \label{item:1}$\sigma \in M$, and
  \item \label{item:2} $\val(G,\sigma) = \{ \val(G,\tau) , \val(G,\rho) \}$
\end{enumerate}

By the implementation of pairs  in $\ZFC$, showing (\ref{item:1})
involves using that the
pairing axiom holds in $M$ and the absoluteness of pairing
thanks to $M$ being transitive.

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ pairs{\isacharunderscore}in{\isacharunderscore}M\ {\isacharcolon}\ \isanewline
\ \ {\isachardoublequoteopen}\ {\isasymlbrakk}\ upair{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharparenright}\ {\isacharsemicolon}\ a\ {\isasymin}\ M\ {\isacharsemicolon}\ b\ {\isasymin}\ M\ {\isacharsemicolon}\ c\ {\isasymin}\ M\ {\isacharsemicolon}\ d\ {\isasymin}\ M\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\isasymlangle}a{\isacharcomma}c{\isasymrangle}{\isacharcomma}{\isasymlangle}b{\isacharcomma}d{\isasymrangle}{\isacharbraceright}\ {\isasymin}\ M{\isachardoublequoteclose}
\end{isabelle}

Item (\ref{item:1}) then follows because \isa{\isasymtau}, \isa{\isasymrho} and
\isa{one} belong to \isa{M} (the last fact holds because \isa{one\isasymin P}, \isa{P\isasymin M} and
\isa{M} is transitive).

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ sigma{\isacharunderscore}in{\isacharunderscore}M\ {\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}upair{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharparenright}\ {\isasymLongrightarrow}\ one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ {\isasymtau}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isasymrho}\ {\isasymin}\ M\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\isasymlangle}{\isasymtau}{\isacharcomma}one{\isasymrangle}{\isacharcomma}{\isasymlangle}{\isasymrho}{\isacharcomma}one{\isasymrangle}{\isacharbraceright}\ {\isasymin}\ M{\isachardoublequoteclose}

\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ pairs{\isacharunderscore}in{\isacharunderscore}M{\isacharcomma}simp{\isacharunderscore}all\ add{\isacharcolon}\ upair{\isacharunderscore}ax{\isacharunderscore}def\ one{\isacharunderscore}in{\isacharunderscore}M{\isacharparenright}%
\end{isabelle}

Under the assumption that \isa{one} belongs to the set \isa{G},
(\ref{item:2}) follows from \isa{def\_val} almost automatically:

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ valsigma\ {\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}one\ {\isasymin}\ G\ {\isasymLongrightarrow}\ {\isacharbraceleft}{\isasymlangle}{\isasymtau}{\isacharcomma}one{\isasymrangle}{\isacharcomma}{\isasymlangle}{\isasymrho}{\isacharcomma}one{\isasymrangle}{\isacharbraceright}\ {\isasymin}\ M\ {\isasymLongrightarrow}\isanewline
\ \ \ val{\isacharparenleft}G{\isacharcomma}{\isacharbraceleft}{\isasymlangle}{\isasymtau}{\isacharcomma}one{\isasymrangle}{\isacharcomma}{\isasymlangle}{\isasymrho}{\isacharcomma}one{\isasymrangle}{\isacharbraceright}{\isacharparenright}\ {\isacharequal}\ {\isacharbraceleft}val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}{\isacharcomma}val{\isacharparenleft}G{\isacharcomma}{\isasymrho}{\isacharparenright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

This ends the proof of the Pairing Axiom in $M[G]$.