### 2018-11-17 LSFA/cover_letter.tex
descripcion de cambios

\documentclass[12pt,english]{article}
\usepackage{babel}
%
\input{header-draft}
%
%
\begin{document}
\title{First steps towards a formalization of Forcing
  \\ Revisions made to the paper}
\author{E.~Gunther \and  M.~Pagano \and P.~S\'anchez~Terraf
  %% fourth author was partially
  %% supported by CONICET, ANPCyT project PICT 2012-1823, and by
  %% SeCyT-UNC project 05/B284.}
}
\date{}
\maketitle

In first place we want to thank the referee for reading the paper very
carefully and for his detailed and constructive criticism. Below we
detail the main changes that have been made%; also, a diff-PDF is
%provided that shows all the changes in color
.

We have spelled-check the manuscript, corrected punctuation, and took
notice of all the typos indicated in the report.

The Introduction has many changes. A whole new section (``Related
work'') has been added, where the formalization of mathematics in
general and of forcing and set theory in particular are discussed.
We also remark that our  development follows the
presentation by Kunen \cite{kunen2011set}.

%
%
\bibliographystyle{mi-estilo-else}
\bibliography{citados}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: