### 2018-11-19 LSFA/names.tex

\subsection{Names}
\label{sec:names}
We formalize the function $\val$ that allows to
construct the elements of the generic extension $M[G]$ from elements
of the ctm $M$ and the generic filter $G$. The definition of $\val$
can be written succintly as a recursive equation
%
\begin{equation}\label{eq:def-val}
\val(G,\tau)\defi \{\val(G,\sigma) : \exists p\in \PP.\,
(\lb\sigma,p\rb\in\tau \y p\in G)\}.
\end{equation}
%
The justification that $\val$ is well-defined comes from a general
result (transfinite recursion on well-founded
relations~\cite[p. 48]{kunen2011set}). Given a well-founded relation
$R \subseteq A \times A$ and a functional
$H : A \times (A \to A) \to A$, the principle asserts the existence of
a function $F : A \to A$ satisfying
$F(a) = H(a,F\uparrow\,(R^{-1}(a)))$. This principle is formalized in
Isabelle/ZF and one can use the operator \isa{wfrec}\footnote{Notice
  that this form of recursive definitions is more general than the one
  used in the previous section to define
  \isa{dc{\isacharunderscore}witness}.} to define functions using
transfinite recursion. To be precise, \isa{wfrec :: [i, i, [i,i]=>i]
  => i} is a slight variation, where the first argument is the
relation, the third is the functional, and the second corresponds to
the argument of $F$. Notice that the relation and the function argument
of the functional are internalized as terms of type \isa{i}.

In our case the functional is called $Hv$ and takes an additional argument for the
parameter $G$:
\[
  \mathit{Hv}(G,y,f) = \{f(x) : x \in dom(y) \y \exists p\in \PP.\,
(\lb x,p\rb\in y \y p\in G)\}
\]
while the relation is given by:
\[
x \mathrel{\mathit{ed}} y \iff \exists p . \lb x,p\rb\in y.
\]
Recall that in $\ZFC$, an ordered pair $\lb x,y \rb$ is the set
$\{\{x\},\{x,y\}\}$. It is trivial to deduce the well-foundedness of
$\mathit{ed}$ from the fact that $\in$ is well-founded, which follows
from the Foundation Axiom.

In our formalization of this recursion,  the first argument of
\isa{wfrec} is the term of type  \isa{i} obtained by restricting the
relation $\mathit{ed}$ to  a set:
% 
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ edrel\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}edrel{\isacharparenleft}A{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}{\isacharless}x{\isacharcomma}y{\isachargreater}\ {\isasymin}\ A{\isacharasterisk}A\ {\isachardot}\ x\ {\isasymin}\ domain{\isacharparenleft}y{\isacharparenright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}
%
Since \isa{edrel(A)} is a subset of a  well-founded relation (the
transitive closure of the membership relation restricted to \isa{A}),
then it is well-founded as well.

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ wf{\isacharunderscore}edrel\ {\isacharcolon}\ {\isachardoublequoteopen}wf{\isacharparenleft}edrel{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ wf{\isacharunderscore}subset\ {\isacharbrackleft}of\ {\isachardoublequoteopen}trancl{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}edrel{\isacharunderscore}sub{\isacharunderscore}memrel\ wf{\isacharunderscore}trancl\ wf{\isacharunderscore}Memrel{\isacharparenright}\isanewline
\ \ \isacommand{done}\isamarkupfalse%
%
\end{isabelle}
%
All but one lemma used in the above proof
(\isa{wf{\isacharunderscore}subset},
\isa{wf{\isacharunderscore}trancl}, and
\isa{wf{\isacharunderscore}Memrel}) are already
present in Isabelle/ZF. The remaining technical result has
been proved using the Isar language:
%
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ edrel{\isacharunderscore}sub{\isacharunderscore}memrel{\isacharcolon}\ {\isachardoublequoteopen}edrel{\isacharparenleft}A{\isacharparenright}\ {\isasymsubseteq}\ trancl{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}A{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

The formalizaton of the functional $\mathit{Hv}$ is straightforward and $\val$ is defined using \isa{wfrec}:
%
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ Hv\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hv{\isacharparenleft}G{\isacharcomma}y{\isacharcomma}f{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}\ f{\isacharbackquote}x\ {\isachardot}{\isachardot}\ x{\isasymin}\ domain{\isacharparenleft}y{\isacharparenright}{\isacharcomma}\ {\isasymexists}p{\isasymin}P{\isachardot}\ {\isacharless}x{\isacharcomma}p{\isachargreater}\ {\isasymin}\ y\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ val\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}\ {\isacharequal}{\isacharequal}\ wfrec{\isacharparenleft}edrel{\isacharparenleft}eclose{\isacharparenleft}M{\isacharparenright}{\isacharparenright}{\isacharcomma}\ {\isasymtau}{\isacharcomma}\ Hv{\isacharparenleft}G{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
Then we can recover the recursive expression~(\ref{eq:def-val}) thanks to the
following lemma:
%
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ def{\isacharunderscore}val{\isacharcolon}
\isanewline
\ \ \ \ \ \ \ \  {\isachardoublequoteopen}x{\isasymin}M\ {\isasymLongrightarrow}\ val{\isacharparenleft}G{\isacharcomma}x{\isacharparenright}\ {\isacharequal}\ {\isacharbraceleft}val{\isacharparenleft}G{\isacharcomma}t{\isacharparenright}\ {\isachardot}{\isachardot}\ t{\isasymin}domain{\isacharparenleft}x{\isacharparenright}\ {\isacharcomma}\ {\isasymexists}p{\isasymin}P\ {\isachardot}\ \ {\isasymlangle}t{\isacharcomma}\ p{\isasymrangle}{\isasymin}x\ {\isasymand}\ p\ {\isasymin}\ G\ {\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

We can finally define the generic extension of $M$ by $G$, also
setting up the notation $M[G]$ for it:
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ GenExt\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}i{\isachardoublequoteclose}\ \ \ \ \ {\isacharparenleft}{\isachardoublequoteopen}M{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where} \isanewline
\ \ {\isachardoublequoteopen}GenExt{\isacharparenleft}G{\isacharparenright}{\isacharequal}{\isacharequal}\ {\isacharbraceleft}val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}{\isachardot}\ {\isasymtau}\ {\isasymin}\ M{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}
It is conventional in Isabelle/ZF to define introduction and
destruction rules for definitions like \isa{GenExt}; in our case, it is
enough to know $x\in M$ in order to know $\val(G,x) \in M[G]$:
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ GenExtI{\isacharcolon}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ val{\isacharparenleft}G{\isacharcomma}x{\isacharparenright}\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}
\end{isabelle}
\noindent The destruction rule corresponding to the generic extension says
that any $x \in M[G]$ comes from some $\tau \in M$ via $\val$. 
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ GenExtD{\isacharcolon}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}\ {\isasymLongrightarrow}\ {\isasymexists}{\isasymtau}{\isasymin}M{\isachardot}\ x\ {\isacharequal}\ val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

%% It is to be noted that this results holds under the assumption that
%% the name \isa{t} is in $M$.

We now provide names for elements in $M$. That is, for each $x\in M$,
we define $\chk(x)$ (usually denoted by $\check{x}$ in the literature)
such that $\val(G,\chk(x))=x$. This will show that $M\sbq M[G]$, with
a caveat we make explicit in the end of this section. As explained in
the introduction, the fact that $M[G]$ extends $M$ is crucial to show
that $\ZFC$ holds in the former.
%
The definition of $\chk(x)$ is a straightforward $\in$-recursion:
\begin{equation}
  \label{eq:def-check}
  \chk(x)\defi\{\lb\chk(y),\1\rb : y\in x\}
\end{equation}
Now the set-relation argument for \isa{wfrec} is the membership
relation restricted to a set \isa{A}, \isa{Memrel(A)}.

\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\ \isanewline
\ \ Hcheck\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}i{\isacharcomma}i{\isacharbrackright}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Hcheck{\isacharparenleft}z{\isacharcomma}f{\isacharparenright}\ \ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}\ {\isacharless}f{\isacharbackquote}y{\isacharcomma}one{\isachargreater}\ {\isachardot}\ y\ {\isasymin}\ z{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ check\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}check{\isacharparenleft}x{\isacharparenright}\ {\isacharequal}{\isacharequal}\ wfrec{\isacharparenleft}Memrel{\isacharparenleft}eclose{\isacharparenleft}{\isacharbraceleft}x{\isacharbraceright}{\isacharparenright}{\isacharparenright}{\isacharcomma}\ x\ {\isacharcomma}\ Hcheck{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
Here, \isa{eclose} returns the (downward) $\in$-closure of its
argument. The main lemmas about $\val$ and $\chk$ require some
instances of replacement for $M$; we set up a locale to assemble these
assumptions:
\begin{isabelle}
\isacommand{locale}\isamarkupfalse%
\ M{\isacharunderscore}extra{\isacharunderscore}assms\ {\isacharequal}\ forcing{\isacharunderscore}data\ {\isacharplus}\isanewline
\ \ \isakeyword{assumes}\ check{\isacharunderscore}in{\isacharunderscore}M\ {\isacharcolon}\ \ \ \ \ \  {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ x\ {\isasymin}\ M\ {\isasymLongrightarrow}\ check{\isacharparenleft}x{\isacharparenright}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isakeyword{and}\ sats{\isacharunderscore}upair{\isacharunderscore}ax{\isacharcolon}\ \ \ \ {\isachardoublequoteopen}upair{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isakeyword{and}\ repl{\isacharunderscore}check{\isacharunderscore}pair\ {\isacharcolon}\ {\isachardoublequoteopen}strong{\isacharunderscore}replacement{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharcomma}{\isasymlambda}p\ y{\isachardot}\ y\ {\isacharequal}{\isacharless}check{\isacharparenleft}p{\isacharparenright}{\isacharcomma}p{\isachargreater}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

The first assumption asserts that all the relevant names are indeed in
$M$ (i.e., $\chk(x) \in M$ if $x\in M$) and it is needed to prove that
$\val(G,\chk(x))=x$. It will take a serious effort to fulfill this
assumption: One of the hardest parts of Paulson's formalization of
constructibility involves showing that models are closed under
recursive construction. We will eventually formalize that if
$M\models\ZFC$ and the arguments of \isa{wfrec} are in $M$, then its
value also is. This will require to adapt to ctm models several
locales defined in \cite{paulson_2003} that were intended to be used
for the class of constructible sets. Notice that the only requirement
on the set \isa{G} is that it contains the top element of the poset
\isa{P}.

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ valcheck\ {\isacharcolon}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}one\ {\isasymin}\ G{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}y\ {\isasymin}\ M\ {\isasymLongrightarrow}\ val{\isacharparenleft}G{\isacharcomma}check{\isacharparenleft}y{\isacharparenright}{\isacharparenright}\ \ {\isacharequal}\ y{\isachardoublequoteclose}
\end{isabelle}

\subsection{Basic results about the generic extension}

We turn now to prove that $M[G]$ is transitive and $G\in
M[G]$. Showing that $M[G]$ is transitive amounts to prove $y \in M[G]$
for any $x \in M[G]$ and $y \in x$. 
% \begin{isabelle}
% \isacommand{definition}\isamarkupfalse%
% \isanewline
% \ \ Transset\ \ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isacharequal}{\isachargreater}o{\isachardoublequoteclose}\ \ \isakeyword{where}\isanewline
% \ \ \ \ {\isachardoublequoteopen}Transset{\isacharparenleft}i{\isacharparenright}\ {\isacharequal}{\isacharequal}\ {\isasymforall}x{\isasymin}i{\isachardot}\ x{\isacharless}{\isacharequal}i{\isachardoublequoteclose}
% \end{isabelle}

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ trans{\isacharunderscore}Gen{\isacharunderscore}Ext{\isacharprime}\ {\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ \ {\isachardoublequoteopen}x\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}\ \ \isakeyword{and}\ \ {\isachardoublequoteopen}y\ {\isasymin}\ x{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ \ \ \ {\isachardoublequoteopen}y\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}
\end{isabelle}
The proof of this lemma is straightforward because from $x \in M[G]$
we can obtain $\tau \in M$ such that $x = \val(G,\tau)$. Notice also
that using the characterization of $\val$ given by
\isa{def{\isacharunderscore}val} we can extract some
$\theta \in \dom(\tau)$ such that $y =\val(G,\theta)$; to conclude
$\val(G,\theta) \in M[G]$ it is enough to prove $\theta \in M$, which
follows from the transitivity of $M$.

In contrast, the proof that $G\in M[G]$ is more demanding. In fact, we
set $\dot{G} = \{ \langle \check{p},p\rangle \,|\, p \in P \}$ as a
putative name for $G$. Proving that $\dot{G}$ is in fact a name for
$G$ requires to prove that $\dot{G} \in M$, using an instance of
replacement for $M$ (\noindent namely that given by the assumption
\isa{repl{\isacharunderscore}check{\isacharunderscore}pair}), and
then proving that $\val(G,\dot{G})=G$.

\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\isanewline
\ \ G{\isacharunderscore}dot\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}G{\isacharunderscore}dot\ {\isacharequal}{\isacharequal}\ {\isacharbraceleft}{\isacharless}check{\isacharparenleft}p{\isacharparenright}{\isacharcomma}p{\isachargreater}\ {\isachardot}\ p{\isasymin}P{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ G{\isacharunderscore}dot{\isacharunderscore}in{\isacharunderscore}M\ {\isacharcolon} {\isachardoublequoteopen}G{\isacharunderscore}dot\ {\isasymin}\ M{\isachardoublequoteclose}
\isanewline\isanewline
\isacommand{lemma}\isamarkupfalse%
\ val{\isacharunderscore}G{\isacharunderscore}dot\ {\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}G\ {\isasymsubseteq}\ P{\isachardoublequoteclose}\ \ \isakeyword{and}\ \ 
{\isachardoublequoteopen}one\ {\isasymin}\ G{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}val{\isacharparenleft}G{\isacharcomma}G{\isacharunderscore}dot{\isacharparenright}\ {\isacharequal}\ G{\isachardoublequoteclose}
\end{isabelle}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% TeX-master: "first_steps_into_forcing"
%%% End: