### 2018-11-30 LICS/intro.tex
modularizando el manuscrito

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Introduction}
% no \IEEEPARstart
We follow \cite{kunen2011set}. An introduction to the issues discussed
in this paper can be found in \cite{2018arXiv180705174G}. An
introduction to forcing can be found at \cite{chow-beginner-forcing},
and the book \cite{weaver2014forcing} contains a thorough treament
minimizing the technicalities.

An enumeration of achievements to be presented:
\begin{enumerate}
\item Discussion of the Isabelle view of life.
  \begin{enumerate}
  \item Meta (Pure) vs Object (FOL+ZF) theory.
  \item Object theory is presented axiomatically; in particular,
    first-order ``formulas'' are not such (i.e., the \emph{initial}
    algebra of formulas is \textbf{not} being specified). Thus, we are
    not able to perform recursion on terms of type \tyo.
  \end{enumerate}
\item Discussion of locales.
  \begin{enumerate}
  \item How they help presenting complex results.
  \item Their key role in stating the Fundamental Theorems (without
    defining \isatt{forces}).
  \item \isatt{forces} is defined for elements of \formula.
  \end{enumerate}
\item Enhancements in the recursion machinery.
  \begin{enumerate}
  \item Impact on names.
  \end{enumerate}
\item Hacking of \isatt{ZF-Constructible}.
  \begin{enumerate}
  \item Elimination of the \tyo-schematic Replacement Axiom from
    locales.
  \item Replacing \isatt{M(nat)} by \isatt{M(0)} (merely non
    emptiness) from Relative (and other places?).
  \end{enumerate}
\item Inteface between first-order properties of $M$ and those of
  $\#\#M$.
  \begin{enumerate}
  \item Enumerate properties and statistics.
  \end{enumerate}
\item Details of the proof of Separation.
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% End: