### 2018-12-02 LICS/proof-of-separation.tex
trayendo cosas de blueprint sin ningún reparo

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Proof of Separation}

This proof can be found in the file \verb|Separation_Axiom.thy| of the
development, which we proceed to discuss.

The key technical result is the following:
\begin{isabelle}
  \isacommand{lemma}\isamarkupfalse%
  \ Collect{\isacharunderscore}sats{\isacharunderscore}in{\isacharunderscore}MG\ {\isacharcolon}\isanewline
  \ \ \isakeyword{assumes}\isanewline
  \ \ \ \ {\isachardoublequoteopen}{\isasympi}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymsigma}\ {\isasymin}\ M{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}G{\isacharcomma}\ {\isasympi}{\isacharparenright}\ {\isacharequal}\ c{\isachardoublequoteclose}\ {\isachardoublequoteopen}val{\isacharparenleft}G{\isacharcomma}\ {\isasymsigma}{\isacharparenright}\ {\isacharequal}\ w{\isachardoublequoteclose}\isanewline
  \ \ \ \ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\ {\isachardoublequoteopen}arity{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymle}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
  \ \ \isakeyword{shows}\ \ \ \ \isanewline
  \ \ \ \ {\isachardoublequoteopen}{\isacharbraceleft}x{\isasymin}c{\isachardot}\ sats{\isacharparenleft}M{\isacharbrackleft}G{\isacharbrackright}{\isacharcomma}\ {\isasymphi}{\isacharcomma}\ {\isacharbrackleft}x{\isacharcomma}\ w{\isacharbrackright}{\isacharparenright}{\isacharbraceright}{\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}
\end{isabelle}
%
From this, using absoluteness, we will be able to derive the
$\phi$-instance of Separation.

To show that   
\[
S\defi\{x\in c : M[G],[x,w]\models \phi(x_0,x_1)\} \in M[G],
\]
it is enough to provide a name $n\in M$ for this set.

The candidate name is
\[
n \defi \{u \in\dom(\pi)\times\PP :M,[u,\PP,\leq,\1,\sig,\pi]\models \psi\}
\]
where
\[
\psi \defi \exists \th\, p.\ x_0=\lb\th,p\rb \y 
   \forceisa(\th\in x_5\y\phi(\th,x_4)).
\]
The fact that $n\in M$ follows by an application of a six-variable
instance of Separation in $M$ (lemma \isatt{six{\isacharunderscore}sep{\isacharunderscore}aux}).

Almost a third part of the proof involves the syntactic handling of
internalized formulas and permutation of variables. The more
substantive portion concerns proving that actually $\val(G,n)=S$.

Let's first focus into the predicate 
\begin{equation}\label{eq:1}
M,[u,\PP,\leq,\1,\sig,\pi]\models \psi
\end{equation}
defining $n$ by separation. By definition of the satisfaction
relation and %% permuting variables
absoluteness, we have that it is equivalent to the fact
that there exist $\th,p\in M$ with   $u=\lb\th,p\rb$  and 
\[
M,[\PP,\leq,\1,p,\th,\sig,\pi]\models \forceisa(x_4\in
x_6\y\phi(x_4,x_5)). 
\]
% (Note that the variable $x_7$ is not used.)
This, in turn, is equivalent by the Definition of Forcing to: \emph{For all $M$-generic
filters $F$ such that $p\in F$,} 
\begin{equation}\label{eq:2}
M[F],[\val(F,\th),\val(F,\sig),\val(F,\pi)]\models x_0\in
x_2\y\phi(x_0,x_1). 
\end{equation}
We can instantiate this statement with $G$ and obtain
\[
p\in G \impl M[G],[\val(G,\th),w,c]\models x_0\in
x_2\y\phi(x_0,x_1). 
\] 
Let $Q(\th,p)$ be the last statement. We have just seen that
(\ref{eq:1}) implies 
\[
\exists \th,p\in M.\ u=\lb\th,p\rb \y Q(\th,p).
\]
Hence $n$ is included in 
\[
m\defi \{u \in\dom(\pi)\times\PP : \exists \th,p\in M.\ u=\lb\th,p\rb
\y Q(\th,p)\}. 
\]

It can be seen that
\begin{lemma}
  $\val(G,m) = S$.
\end{lemma}
\noindent And by monotonicity of $\val$ we obtain
\begin{lemma}
  $\val(G,n)\sbq \val(G,m)$.
\end{lemma}
To complete the proof, it is therefore enough to show that
$S\sbq \val(G,n)$. For this, let $x\in S$. Hence there exists
$\lb\th,q\rb\in\pi$ such that 
$q\in G$ and $x=\val(G,\th)$.

On the other hand, since 
\[
M[G],[\val(G,\th),\val(G,\sig),\val(G,\pi)]\models
 x_0\in x_2\y\phi( x_0, x_1),
\]
by the  Truth Lemma there must exist $r\in G$ such that
\[
M,[\PP,\leq,\1,r,\th,\sig,\pi]\models
\forceisa(x_4\in x_6\y\phi( x_4, x_5)).
\]
Since $G$ is a filter, there is $p\in G$ such that $p\leq q, r$.
By Strenghtening, we have
\[
M,[\PP,\leq,\1,p,\th,\sig,\pi]\models
\forceisa(x_4\in x_6\y \phi( x_4, x_5)),
\]
which by the Truth Lemma gives us: \emph{for all $M$-generic $F$,
  $p\in F$ implies} 
\[
M[F],[\val(F,\th),\val(F,\sig),\val(F,\pi)]\models
 x_0\in  x_2 \y\phi( x_0, x_1).
\]
Note this is the same as (\ref{eq:2}). In fact, we have that
\begin{lemma}
  $\val(G,n)$ equals
  \[
  \{\val(G,\th): \th\in\dom(\pi) \y  p \in G 
  \y \text{(\ref{eq:2}) holds}\}.
  \]
\end{lemma}
This shows that $S\sbq \val(G,n)$, since $\th\in\dom(\pi)$ by construction.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: