### 2018-12-04 LICS/generic-extensions.tex
cambio esa horrible font tt y agrego secciones

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Generic extensions}

We will swiftly review some definitions in order to reach the concept
of \emph{generic extension}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: