### 2018-12-07 LICS/renaming.tex
Más texto en renaming.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Renaming}
\newcommand{\renaming}[2]{(#1)[#2]}
\newcommand{\inFm}[2]{#1 \in #2}
\newcommand{\eqFm}[2]{#1 = #2}
\newcommand{\negFm}[1]{\neg #1}
\newcommand{\andFm}[2]{#1 \wedge #2}
\newcommand{\forallFm}[1]{\forall #1}

\newcommand{\inIFm}[2]{\mathsf{Member}(#1,#2)}
\newcommand{\eqIFm}[2]{\mathsf{Equal}(#1,#2)}
\newcommand{\nandIFm}[2]{\mathsf{Nand}(#1,#2)}
\newcommand{\forallIFm}[1]{\mathsf{Forall(#1)}}

In the course of our work we need to reason about renaming of formulas
and its effect on their satisfiability. Internalized formulas are
implemented using de Bruijn indices for variables and the arity of a
formula $\phi$ gives the least natural number containing all the free
variables in $\phi$. Following \citet{fiore-abssyn}, one can
understand the arity of a formula as the context of the free
variables; notice that the arity of $\forallFm{\phi}$ is the
predecessor of the arity of $\phi$. Renamings are, consequently,
mappings between finite sets; since we can think of $\mathsf{succ}(n)$
as the co-product $1+n$, then given a renaming $f \colon n \to m$, the
unique morphism $\mathsf{id}_1+f \colon 1+n \to 1+m$ is used to rename
free variables in a quantified formula.

\begin{definition}[Renaming]
  Let $\phi$ be a formula of arity $n$ and let $f \colon n \to m$, the
  renaming of $\phi$ by $f$, denoted $\renaming{\phi}{f}$, is defined
  by recursion on $\phi$:
  \begin{gather*}
    \renaming{\inFm{i}{j}}{f} = \inFm{f\,i}{f\,j}\\
    \renaming{\eqFm{i}{j}}{f} = \eqFm{f\,i}{f\,j}\\
    \renaming{\negFm{\phi}}{f} = \negFm{\renaming{\phi}{f}}\\
    \renaming{\andFm{\phi}{\psi}}{f} = \andFm{\renaming{\phi}{f}}{\renaming{\psi}{f}}\\
    \renaming{\forallFm{\phi}}{f} = \forallFm{\renaming{\phi}{\mathsf{id}+f}}
  \end{gather*}
\end{definition}

Environments can be thought, then, as mappings from a finite set to
the model $M$; the action of renaming on environments is simply given
by composition. An easy proof connects satisfaction with renamings.
\begin{lemma}
  \label{lem:renaming}
  Let $\phi$ be a formula of arity $n$, $f \colon n \to m$ be a
  renaming, and let $\rho \colon n \to M$ and $\rho' \colon m \to M$
  be environments. If for all $i \in n$,
  $\rho\,i = (\rho' \circ f)\,i)$, then $M,\rho\models \phi$ is
  equivalent to $M,\rho' \models \renaming{\phi}{f}$.
\end{lemma}

An important resource in Isabelle/ZF is the facility for defining
inductive sets \cite{paulson2000fixedpoint,paulson1995set} together
with a principle for defining functions by structural recursion.
Internalized formulas are a prime example of this, so we define
a function \isa{ren} that associates to each formula an internalized
function that can be later applied to suitable arguments.
\begin{isabelle}
  \ \isacommand{consts}\isamarkupfalse%
\ \ \ ren\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isacharequal}{\isachargreater}i{\isachardoublequoteclose}\isanewline
\ \isacommand{primrec}\isamarkupfalse%
\isanewline
\ {\isachardoublequoteopen}ren{\isacharparenleft}Member{\isacharparenleft}x{\isacharcomma}y{\isacharparenright}{\isacharparenright}\ {\isacharequal}\isanewline
\ \ {\isacharparenleft}{\isasymlambda}n{\isasymin}nat{\isachardot}{\isasymlambda}f\ {\isasymin}\ bij{\isacharparenleft}n{\isacharcomma}n{\isacharparenright}{\isachardot}\ Member\ {\isacharparenleft}f{\isacharbackquote}x{\isacharcomma}\ f{\isacharbackquote}y{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ {\isachardoublequoteopen}ren{\isacharparenleft}Equal{\isacharparenleft}x{\isacharcomma}y{\isacharparenright}{\isacharparenright}\ {\isacharequal}\isanewline
\ \ {\isacharparenleft}{\isasymlambda}n{\isasymin}nat{\isachardot}\ {\isasymlambda}f\ {\isasymin}\ bij{\isacharparenleft}n{\isacharcomma}n{\isacharparenright}{\isachardot}\ Equal\ {\isacharparenleft}f{\isacharbackquote}x{\isacharcomma}\ f{\isacharbackquote}y{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \isanewline
\ {\isachardoublequoteopen}ren{\isacharparenleft}Nand{\isacharparenleft}p{\isacharcomma}q{\isacharparenright}{\isacharparenright}\ {\isacharequal}\isanewline
\ \ {\isacharparenleft}{\isasymlambda}n{\isasymin}nat{\isachardot}\ {\isasymlambda}f\ {\isasymin}\ bij{\isacharparenleft}n{\isacharcomma}n{\isacharparenright}{\isachardot}\isanewline
\ \ \ \ \ \ \ \ Nand\ {\isacharparenleft}ren{\isacharparenleft}p{\isacharparenright}{\isacharbackquote}n{\isacharbackquote}f{\isacharcomma}\ ren{\isacharparenleft}q{\isacharparenright}{\isacharbackquote}n{\isacharbackquote}f{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \isanewline
\ {\isachardoublequoteopen}ren{\isacharparenleft}Forall{\isacharparenleft}p{\isacharparenright}{\isacharparenright}\ {\isacharequal}\isanewline
\ \ {\isacharparenleft}{\isasymlambda}n{\isasymin}nat{\isachardot}\ {\isasymlambda}f\ {\isasymin}\ bij{\isacharparenleft}n{\isacharcomma}n{\isacharparenright}{\isachardot}\isanewline
\ \ \ \ \ \ \ \ Forall\ {\isacharparenleft}ren{\isacharparenleft}p{\isacharparenright}\ {\isacharbackquote}succ{\isacharparenleft}n{\isacharparenright}{\isacharbackquote}\ sum{\isacharunderscore}id{\isacharparenleft}n{\isacharcomma}f{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

Notice that our implementation of renaming is more restricted as we
work with bijections; \isa{sum{\isacharunderscore}id} corresponds to
the co-product morfism $\mathsf{id}_{1}+f \colon 1 + n \to 1 +
n$. This definition of renaming exposes some inconveniences of working
in the untyped realm of set theory, for example to use \isa{ren} we
will need to prove that the renaming is a bijection. Another
inconvenience is that environments are not finite mappings but lists
(yet another inductively defined set); given an environment and a
bijection, the following definition returns the composition
$\rho \circ f$ as a list:
\begin{isabelle}
  \isacommand{definition}\isamarkupfalse%
\ perm{\isacharunderscore}list\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}i{\isacharcomma}i{\isacharbrackright}\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ {\isachardoublequoteopen}perm{\isacharunderscore}list{\isacharparenleft}f{\isacharcomma}\isasymrho{\isacharparenright}\ {\isacharequal}{\isacharequal}tab{\isacharparenleft}nth{\isacharunderscore}i{\isacharparenleft}\isasymrho{\isacharparenright}\ O\ f{\isacharcomma}length{\isacharparenleft}\isasymrho{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
In the previous definition \isa{tab(g,n)} creates a list tabulating the
image of the function \isa{g} on the elements of \isa{n}. 
% \begin{isabelle}
% \isacommand{lemma}\isamarkupfalse%
% \ perm{\isacharunderscore}list{\isacharunderscore}tc\ \ {\isacharcolon}\ {\isachardoublequoteopen}\ \isasymrho\ {\isasymin}\ list{\isacharparenleft}A{\isacharparenright}\ {\isasymLongrightarrow}\ \isanewline
% \ \ \ \ f{\isasymin}\ bij{\isacharparenleft}length{\isacharparenleft}\isasymrho{\isacharparenright}{\isacharcomma}length{\isacharparenleft}\isasymrho{\isacharparenright}{\isacharparenright}\ {\isasymLongrightarrow}\ \isanewline
% \ \ \ \ perm{\isacharunderscore}list{\isacharparenleft}f{\isacharcomma}\isasymrho{\isacharparenright}\ {\isasymin}\ list{\isacharparenleft}A{\isacharparenright}{\isachardoublequoteclose}
% \end{isabelle}
The result connecting renaming with satisfiability, as in
\ref{lem:renaming}, is the following:
\fbox{Miguel: esto hay que acomodarlo o no mostrarlo!}
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ renSat\ {\isacharcolon}\ \isanewline
\ \ \isakeyword{fixes}\ {\isachardoublequoteopen}{\isasymphi}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymin}\ formula{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ \ {\isachardoublequoteopen}{\isasymAnd}\ {\isasymrho}\ f\ {\isachardot}\ {\isasymlbrakk}\ {\isasymrho}\ {\isasymin}\ list{\isacharparenleft}M{\isacharparenright}\ {\isacharsemicolon}\ f\ {\isasymin}\ bij{\isacharparenleft}length{\isacharparenleft}{\isasymrho}{\isacharparenright}{\isacharcomma}length{\isacharparenleft}{\isasymrho}{\isacharparenright}{\isacharparenright}\ {\isacharsemicolon}\ arity{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymle}\ length{\isacharparenleft}{\isasymrho}{\isacharparenright}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ sats{\isacharparenleft}M{\isacharcomma}{\isasymphi}{\isacharcomma}{\isasymrho}{\isacharparenright}\ {\isasymlongleftrightarrow}\ sats{\isacharparenleft}M{\isacharcomma}rename{\isacharparenleft}{\isasymphi}{\isacharparenright}{\isacharbackquote}length{\isacharparenleft}{\isasymrho}{\isacharparenright}{\isacharbackquote}converse{\isacharparenleft}f{\isacharparenright}{\isacharcomma}{\isasymrho}{\isasymcirc}f{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: