### 2018-12-07 LICS/conclusion.tex
ayuda memoria en conclusión

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Conclusions and future work}
The proof of the Powerset Axiom does not present difficulties.

The proof of the Replacement Axiom, both in Kunen and in
other presentations \cite{neeman-course} require a relativized proof
of the Reflection Principle.

\section*{Acknowledgment}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% End: