### 2018-12-14 LICS/easy-axioms.tex
Descripción corta de locales (copiado de Whatsapp)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Foundation, Union, Infinity}
\label{sec:easy-axioms}

It is straightforward to show that the generic extension $M[G]$
satisfies foundation, union, and infinity. The easiest one is
foundation which does not depend on the satisfaction by $M$.

\fbox{Hablar de \texttt{M\_trivial} y del uso de interfaz para instanciar}

\fbox{Falta \texttt{check\_in\_M} para completar Infinity}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: