### 2018-12-17 LICS/hacking.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Hacking of \isatt{ZF-Constructible}}

There are several lemmas that were declared, in later developments, as
introduction/simplification rules (notably, the rule
\isatt{equalityI}). They raised a warning, and we have 
eliminated some of them, but in some cases we had to keep them because
the proof works because it is insisted that the rule is \emph{safe}.

The most notable changes, located in the theory \texttt{Relative}, are
the following:
\begin{enumerate}
\item\label{item:1} We eliminated the requirement that the relative Axiom of Replacement
  is satisfied by a class $M$ to be in the locale \isatt{M\_trivial}. 
\item\label{item:2} We moved the requirement of the Powerset Axiom to \isatt{M\_basic}. 
\item\label{item:3} We replaced the need that the set of natural numbers is in $M$ by the
  milder hypothesis that $M(0)$.
\end{enumerate}

As a consequence of Item~\ref{item:1}, the lemma
\isatt{strong\_replacementI} is no longer valid and was commented
out.

We moved the requirement $M(\mathtt{nat})$ to the locale
\isatt{M\_trancl}, where it is needed for the first time. Some results,
for instance \isatt{rtran\_closure\_mem\_iff} and 
\isatt{iterates\_imp\_wfrec\_replacement} had to be moved inside that
locale.

The proof, for instance, that the constructible universe $L$ satisfies
the modified locale \isatt{M\_trivial} holds with minor
modifications. Nevertheless, in order to have a neater presentation,
we have stripped off several sections concerning $L$ from the theories
\isatt{L\_axioms} and \isatt{Internalize}, and we merged them to form
the new file  \isatt{Internalizations}. Actually we believe that the
\isatt{ZF-Constructible} session  would benefit from some changes in
the way some results are organized accross theory files; for
intance, by cataloging all or most of the internalized formulas in one
file.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: