### 2018-12-18 LICS/easy-axioms.tex
Primer texto para foundation y extensionality

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Extensionality, Foundation, Union, Infinity}
\label{sec:easy-axioms}
\newcommand{\forallRel}[2]{\forall #1\kern -1pt[#2]}
It is straightforward to show that the generic extension $M[G]$
satisfies extensionality, foundation, union, and infinity. % It is
% well-known that extensionality is satisfied by any
% transitive class.\footnote{We plan to define a new hierarchy of
%   locales for ZF where transitive classes will be at the base.}
To say that $A$ satisfies some axiom, say extensionality, means
that the relativized version of extensionality is true in $A$:
\[
\forallRel{x}{A}. \forallRel{y}{A}. (\forallRel{w}{A} . w \in x \leftrightarrow w\in y) \rightarrow x = y
\]

Extensionality for $M[G]$ is straightforward because $M[G]$ is
transitive. %\footnote{We generalized slightly some results by Paulson:
%  in fact, several basic lemmas are valid for any transitive class
%  (for instance, absoluteness of Union). We plan to define a new
%  hierarchy of locales for ZF where transitive classes will be at the
%  base.}
In fact, for $x,y\in M[G]$, we can prove
$\forall w. w\in x \leftrightarrow w\in y$ (notice that the binding
$w$ is not relativized), because $M[G]$ is transitive; therefore by
extensionality we conclude $x=y$.

Foundation for $M[G]$ does not depend on $M[G]$ being transitive: in
this case we take $x\in M[G]$ and prove that there is an
$\in$\kern -1pt-minimal element, relativized to $M[G]$, in $x$. Using
foundation for $x\cap M[G]$ we get a minimal $y$ (which is not
relativized to $M[G]$), so it is still minimal when relativized to
$M[G]$.

The Union axiom asserts that if $x$ is a set, then there exists
another set (the union of $x$) containing all the elements in each
element of $x$. The relativized version of Union asks to give a name
$\pi_a$ for each $a\in M[G]$ and proving $\val(G,\pi)=\bigcup a$.

\fbox{Hablar de \texttt{M\_trivial} y del uso de interfaz para
  instanciar}

\fbox{Falta \texttt{check\_in\_M} para completar Infinity}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: