### 2018-12-21 LICS/easy-axioms.tex
Ligeras variaciones en easy-axioms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Extensionality, Foundation, Union, Infinity}
\label{sec:easy-axioms}

In our first presentation of this project \cite{2018arXiv180705174G},
we proved tha $M[G]$ satisfies Pairing. It is straightforward to show
that the generic extension $M[G]$ satisfies extensionality and
foundation. Showing that it is closed under Union depends on $G$ being
a filter. Infinity is also easy, but it depends in one further
assumption.

For Extensionality in $M[G]$, the assumption 
$\forall w[M[G]]. w\in x \leftrightarrow w\in y$ yields 
$\forall w. w\in x \leftrightarrow w\in y$ by transitivity of $M[G]$. %\footnote{We generalized slightly some results by Paulson:
%  in fact, several basic lemmas are valid for any transitive class
%  (for instance, absoluteness of Union). We plan to define a new
%  hierarchy of locales for ZF where transitive classes will be at the
%  base.}
Therefore, by (ambient) Extensionality we conclude $x=y$.

Foundation for $M[G]$ does not depend on $M[G]$ being transitive: in
this case we take $x\in M[G]$ and prove, relativized to $M[G]$,  that there is an
$\in$\kern -1pt-minimal element in $x$. Instantiating the global Foundation
Axiom for $x\cap M[G]$ we get a minimal $y$, so it is still minimal
when considered relative to $M[G]$.

It is noteworthy that the proofs in the Isar dialect of Isabelle
strictly follow the argumentation of the two previous paragraphs.

The Union Axiom asserts that if $x$ is a set, then there exists
another set (the union of $x$) containing all the elements in each
element of $x$. The relativized version of Union asks to give a name
$\pi_a$ for each $a\in M[G]$ and proving $\val(G,\pi_a)=\union a$.
Let $\tau$ be the name for $a$, i.e.\ $a=\val(G,\tau)$; 
\citet{kunen2011set} gives $\pi_a$ in terms of $\tau$:
\begin{align*}
  \pi_a = \{\langle\theta,p \rangle :  %&\in \dom(\union(\dom(\tau))) \times \PP : \\
\exists \langle\sigma,q\rangle  \in \tau .
 \exists r . \langle \theta,r\rangle \in \sigma \wedge
    p\leqslant r \wedge p \leqslant q \}
\end{align*}
Our formal definition is slightly different in order to ease the proof
of $\pi_a \in M$; as it is defined using Separation, so one needs to
define the domain of separation and also internalize the predicate as
a formula
\isa{union{\isacharunderscore}name{\isacharunderscore}fm}. Instead of
working directly with the internalized formula, we define a predicate
\isa{Unioin{\isacharunderscore}name{\isacharunderscore}body} and prove the equivalence between
\isa{sats(M,union{\isacharunderscore}name{\isacharunderscore}fm,[P,leq,\isasymtau,x])}
and
\isa{Union{\isacharunderscore}Name{\isacharunderscore}body(P,leq,\isasymtau,x)}. The
definition of $\pi_a$ in our formalization is:
\begin{isabelle}
\isacommand{definition}\isamarkupfalse%
\ Union{\isacharunderscore}name\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isasymRightarrow}\ i{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}Union{\isacharunderscore}name{\isacharparenleft}{\isasymtau}{\isacharparenright}\ {\isacharequal}{\isacharequal}\ \isanewline
\ \ \ \ {\isacharbraceleft}u\ {\isasymin}\ domain{\isacharparenleft}{\isasymUnion}{\isacharparenleft}domain{\isacharparenleft}{\isasymtau}{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isasymtimes}\ P\ {\isachardot}\isanewline
\ \ \ \  \ \ \ \ Union{\isacharunderscore}name{\isacharunderscore}body{\isacharparenleft}P{\isacharcomma}leq{\isacharcomma}{\isasymtau}{\isacharcomma}u{\isacharparenright}{\isacharbraceright}{\isachardoublequoteclose}
\end{isabelle}

Once we know $\pi_a \in M$, the equation $\val(G,\pi_a)=\union a$ is
proved by showing the mutual inclusion; in both cases one uses that
$G$ is a filter.
\begin{isabelle}
  \isacommand{lemma}\isamarkupfalse%
\ Union{\isacharunderscore}MG{\isacharunderscore}Eq\ {\isacharcolon}\ \isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}a\ {\isacharequal}\ val{\isacharparenleft}G{\isacharcomma}{\isasymtau}{\isacharparenright}{\isachardoublequoteclose}\ \isakeyword{and}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}filter{\isacharparenleft}G{\isacharparenright}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}{\isasymtau}\ {\isasymin}\ M{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymUnion}\ a\ {\isacharequal}\ val{\isacharparenleft}G{\isacharcomma}Union{\isacharunderscore}name{\isacharparenleft}{\isasymtau}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}
Since Union is absolute for any transitive class we may conclude that
$M[G]$ is closed under Union:
\begin{isabelle}
  \isacommand{lemma}\isamarkupfalse%
  \ union{\isacharunderscore}in{\isacharunderscore}MG\ {\isacharcolon}\ \isanewline
  \ \ \isakeyword{assumes}\ {\isachardoublequoteopen}filter{\isacharparenleft}G{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Union{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharbrackleft}G{\isacharbrackright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

The proof of Infinity for $M[G]$ takes advantage of some absoluteness
results proved in the locale \isa{M{\isacharunderscore}trivial}; this
proof is easy because we assume $\chk(x) \in M$ whenever $x\in
M$. Since we have already proved that $M[G]$ is transitive,
$\emptyset\in M[G]$ assuming $G$ being generic, and also that it
satisfies Pairing and Union, we can instantiate
\isa{M{\isacharunderscore}trivial}:
\begin{isabelle}
\isacommand{sublocale}\isamarkupfalse%
\ G{\isacharunderscore}generic\ {\isasymsubseteq}\ M{\isacharunderscore}trivial{\isachardoublequoteopen}{\isacharhash}{\isacharhash}M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}
\end{isabelle}
We assume that $M$ satisfies Infinity, i.e., that Infinity relativized
to $M$ holds; therefore we obtain $I \in M$ such that $\emptyset\in I$
and, $x \in I$ implies $\mathit{succ}(x)\in I$ by absoluteness of
empty and successor for $M$. Using the assumption that $M$ is closed
under $\chk$, we deduce $\val(G,\chk(I)) = I \in M[G]$.  Now we can
use absoluteness of emptyness and successor, this time for $M[G]$, to
conclude that $M[G]$ satisfies Infinity.
% \begin{isabelle}
%   \isacommand{lemma}\isamarkupfalse%
% \ infinty{\isacharunderscore}in{\isacharunderscore}MG\ {\isacharcolon}\ {\isachardoublequoteopen}infinity{\isacharunderscore}ax{\isacharparenleft}{\isacharhash}{\isacharhash}M{\isacharbrackleft}G{\isacharbrackright}{\isacharparenright}{\isachardoublequoteclose}
% \end{isabelle}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: