### 2018-12-28 LICS/conclusion.tex
empezando a escribir la conclu

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Conclusions and future work}
\label{sec:conclusions-future-work}

The proof that the axioms of $\ZF$ hold in generic extensions is 
the first important landmark in the path of formalization of the
theory of forcing. The task of proving the Fundamental Theorems can be
considered a part of that landmark, but from a workflow point of view
this can be done independently, and one of our design decisions was to
postpone that task.  Next, various forcing notions must be developed
in order to obtain extensions satisfying diverse properties. This
yields relative consistency results.

We first discuss the axioms we are now missing. The proof of the
Powerset Axiom does not present difficulties, and it is one of the
first goals in the immediate future. This is also the case with the
Axiom of Infinity; it can be proved in a direct but cumbersome way. We
preferred to wait to have a full-fledged inteface between ctms and the
locales providing recursive constructions from Paulson's
\isatt{Constructibility} session. Then the current proof will hold
with no extra assumptions. The same goes for the results $M\sbq M[G]$
and $G\in M[G]$.

The Replacement Axiom, however, requires some more work to be
done. Both in Kunen and in other presentations \cite{neeman-course}
it requires a relativized version (i.e., showing that it holds for $M$) of
the \emph{Reflection Principle}. In order to state this meta-theoretic
result by Montague, recall that
an equivalent formulation of the Foundation Axiom states that the 
universe of sets can be decomposed in a transfinite hierarchy of
sets. 
\begin{theorem}
  Let $V_{\al}\defi\union\{\P(V_\be) : \be<\al\}$ for each ordinal
  $\al$. Then each $V_\al$ is a set and 
  $\forall x. \exists\al .\ \Ord(\al) \y x\in V_\al$.  
\end{theorem}
\begin{theorem}[Reflection Principle]\label{th:reflection-principle}
  For every finite $\Phi\sbq\ZF$, $\ZF$ proves: ``There exists
  unboundedly many $\al$ such that $V_\al\models \Phi$.''
\end{theorem}

This result has also other applications; in particular, it dispenses
the need of ctms of the whole of $\ZF$ for consistency proofs (by
repeating  the proof of Lemma~\ref{lem:wf-model-implies-ctm}, now
starting with a  $V_\al$ satisfying $\Phi$).

\medskip
\fbox{Decir que Separaci\'on con 6 variables en $M$}

\fbox{implica Separaci\'on con a lo sumo 2 en $M[G]$}
\medskip

We believe that the
\isatt{ZF-Constructible} session  would benefit from some changes in
the way some results are organized accross theory files; for
intance, by cataloging all or most of the internalized formulas in one
file. Another, most basic example would be to start out with an even
more locale that only assumes $M$ to be a nonempty transitive class,
as many absoluteness results follow from these hypotheses.  
It is our desire to advocate a future work to a thorough
revision of the development of constructibility to maximize modularity.

\section*{Acknowledgment}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: