### 2018-12-28 LICS/intro_m.tex
cosas menores

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Introduction}
% no \IEEEPARstart
Zermelo-Fraenkel Set Theory ($\ZF$) has a prominent place among formal
theories; in particular, it provides a foundation for mathematics and
most of the formal toolkit used everyday by the computer scientist has
also Set Theory at its base (cf.~\cite{paulson1995set}). In this time
of mechanization of mathematics~\cite{avigad2018mechanization}, it
seems natural to ask for a mechanization of the most salient results
of Set Theory.

After G\"odel's Incompleteness Theorems, we cannot expect to have a
formal proof of the consistency of Set Theory in $\ZF$. Besides its own
consistency, there are other results which are undecided by $\ZF$: the
undecidability of Continuum Hypothesis leads to the development of
techniques for independence proofs. First G\"odel introduced the
theory of \emph{inner models}, which gives rise to his model $L$ of
the \emph{Axiom of Constructibility} \cite{godel-L} and proved the
relative consistency of the Axiom of Choice and the Generalized
Continuum Hypothesis with $\ZF$. Thirty years later Paul
J. Cohen~\cite{Cohen-CH-PNAS} devised the technique of \emph{forcing},
which is the only known way of \emph{extending} models of $\ZF$; in
particular, it can be used to prove the relative consistency of the
negation of the Continuum Hypothesis.

In this work we address a substantial part of formalizing the proof
that given a model $M$ of $\ZF$, any \emph{generic extension} $M[G]$
given by forcing also satisfies $\ZF$. As remarked by
\citet[][p.250]{kunen2011set} \enquote{[...] in verifying that $M[G]$
  is a model for set theory, the hardest axiom to verify is
  Comprehension.}  The most important achievement of this paper is the
mechanization in Isabelle of a substantial part of this result; en
route to this, we also formalized the satisfaction by $M[G]$ of
Extensionality, Foundation, Union, and Infinity.

Our development benefited from the extraordinary work done by Lawrence
Paulson \cite{paulson_2003} on the formalization of G\"odel's
constructible universe in the proof assistant \emph{Isabelle}. The
ultimate goal of our project is the formalization of forcing to
complete the mechanization of the independence of the Continuum
Hypothesis. We think that this project constitutes an interesting case
which stresses how feasible is to formally implement mathematics that
involve several leves of reasoning.

The \emph{Formal Abstract} project~\cite{hales-fabstracts} proposes
that the mechanization of mathematics can advance by formalizing some
main result under the explicit assumption of the material
(definitions, propositions, lemmas) on which it is based. In line with
this advice, we profited that the proof that those axioms hold in
generic extension is independent of the \emph{proofs} of the
``fundamental'' theorems of forcing. Let us remark that the definition
of the forces relation is, by itself, quite demanding; the
formalization of it and of the fundamental theorems of forcing roughly
comprises little less than a half of our full project.

It might be a little surprising the lack of formalizations of forcing
and generic extensions. As far as we know, the development of
\citet{JFR6232} in homotopy type theory for constructing generic
extensions in a sheaf-theoretic setting is the unique mechanization of
forcing. This contrast with the fruitful use of forcing techniques to
extend the Curry-Howard isomorphism to classical axioms
\cite{Miquel:2011:FPT:2058525.2059614,lmcs:1070}. Moreover, the
combination of forcing with intuitionistic type theory
\cite{coquand2010note,coquand2012computational} gives rise to an
algorithm to obtain witnesses of the continuity of definable
functionals.

% moreover forcing translations \cite{jaber:hal-01319066}.

\fbox{Parece mucho comienzo sólo para introducir a Kunen}
\fbox{¿lo puedo achurar un poco?}

In a gross simplification, there are two aspects to a formalization
project like this one: thematic and programmatic. The first concerns
the handling of all the theoretical concepts and results in the
subject, while the second involves the practical issues of the
implementation and design. In the case of forcing, the main intricacy
lies in the first aspect. In this sense, following a sensible
presentation of the material is key.  The authoritative reference 
on the subject during the last 30 years has been Kunen's classical
\cite{kunen1980}. In our
formalizaton we have followed a recent rewrite \cite{kunen2011set}
of that  textbook, which presents the material in the same sharp 
style but offering a lot of details. In some sense this project
wouldn't exist without this book. As alternative, introductory
resources, the  interested reader can check
\cite{chow-beginner-forcing}; also, the book \cite{weaver2014forcing}
contains a thorough treament minimizing the technicalities.

We briefly describe the contents of each
section. Section~\ref{sec:isabelle} contains the bare minimun
requirements to understand the (meta)logics used in Isabelle. Next, an
overview of the model theory of set theory is presented in
Section~\ref{sec:axioms-models-set-theory}. There is an ``internal''
representation of first-order formulas as sets, implemented by
Paulson; Section~\ref{sec:renaming} discusses syntactical
transformations of the former, mainly permutation of variables. 
In Section~\ref{sec:generic-extensions} the generic extensions are
succintly reviewed and how the treatment of well founded recursion in
Isabelle was enhanced. We take care of the ``easy axioms'' in
Section~\ref{sec:easy-axioms}; these are the ones that
do not depend on the forcing theorems. We describe the latter in
Section~\ref{sec:forcing}. We adapted the  work by Paulson to our
needs, and this is described in
Section~\ref{sec:hack-constructible}. We present the proof
of the Separation Axiom Scheme in Section~\ref{sec:proof-separation},
which follows closely its implementation. A plan for future work and
some immediate conclusions are offered in
Section~\ref{sec:conclusions-future-work}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: