### 2019-01-05 LICS/separation_arxiv.tex
versión para arXiv

\documentclass{article}
\usepackage{isabelle,isabellesym}
\renewcommand{\ttdefault}{cmtt}
\usepackage{csquotes}
\usepackage{enumitem}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=(\roman*),
}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\input{header-draft}
\makeatletter
\def\foottext{\gdef\@thefnmark{}\@footnotetext}
\makeatother
\newcommand{\keywords}[1]{\foottext{\emph{Keywords:} #1}}
\newcommand{\ack}[1]{\par\bigskip \noindent \emph{Acknowledgment:} #1}

\begin{document}
\title{Separation in Generic Extensions for Isabelle}
\author{Emmanuel Gunther
  \and 
  Miguel Pagano
  \and 
  Pedro S\'anchez Terraf}
\maketitle

\begin{abstract}
  We mechanize, in the proof assistant
  Isabelle, a proof of the
  axiom-scheme of Separation in 
  generic extensions of models of set theory  
  by using the fundamental theorems of forcing.
  We also formalize the satisfaction of the axioms of
  Extensionality, Foundation, Union, and Powerset. The axiom of
  Infinity is likewise treated, under additional assumptions on the ground
  model.
  We also  extend Paulson's library on constructibility  with
  renaming of variables for internalized formulas, an improvement on
  definitions by recursion on well-founded  relations and sharpening
  of the hypotheses in his development of relativization and
  absoluteness.
\end{abstract}
\keywords{
Isabelle/ZF, forcing, names, generic extension, constructibility.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{intro_m}

\input{isabelle}

\input{zfc-axioms}

\input{renaming}

\input{generic-extensions}

\input{recursion}

% \input{prior-results}

\input{hacking}

\input{easy-axioms}

\input{forcing}

\input{proof-of-separation}

\input{conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{mi-estilo-else}
\bibliography{../LSFA/citados}
%\input{separation_arxiv.bbl}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: