### 2019-01-05 LICS/abstract.tex
abstract

\begin{abstract}
  We mechanize, in the proof assistant
  Isabelle, a proof of the
  axiom-scheme of Separation in 
  generic extensions of models of set theory  
  by using the fundamental theorems of forcing.
  We also formalize the satisfaction of the axioms of
  Extensionality, Foundation, Union, and Powerset. The axiom of
  Infinity is likewise treated, under additional assumptions on the ground
  model.
  In order to achieve these goals, we extended Paulson's library on
  constructibility  with 
  renaming of variables for internalized formulas, improved
  results on 
  definitions by recursion on well-founded  relations, 
  and sharpened
  of the hypotheses in his development of relativization and
  absoluteness.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: