### 2019-01-05 LICS/separation_arxiv.tex
titulo e intro

\documentclass{article}
\usepackage{isabelle,isabellesym}
\renewcommand{\ttdefault}{cmtt}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{enumitem}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=(\roman*),
}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\input{header-draft}
\makeatletter
\def\foottext{\gdef\@thefnmark{}\@footnotetext}
\makeatother
\newcommand{\keywords}[1]{\foottext{\emph{Keywords:} #1}}
\newcommand{\ack}[1]{\par\bigskip \noindent \emph{Acknowledgment:} #1}

\hypersetup{
  pdftitle={Mechanization of Separation in Generic Extensions},
  pdfsubject={Computer Science},
  pdfkeywords={Isabelle/ZF, forcing, names, generic extension, constructibility},
  colorlinks,
  linkcolor={blue!40!black},
  citecolor={blue!40!black},
  urlcolor={blue!40!black}
}

\begin{document}
\title{Mechanization of Separation in Generic Extensions}
\author{Emmanuel Gunther
  \and 
  Miguel Pagano
  \and 
  Pedro S\'anchez Terraf}
\maketitle

\input{abstract}
\keywords{
Isabelle/ZF, forcing, names, generic extension, constructibility.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{intro_m}

\input{isabelle}

\input{zfc-axioms}

\input{renaming}

\input{generic-extensions}

\input{recursion}

% \input{prior-results}

\input{hacking}

\input{easy-axioms}

\input{forcing}

\input{proof-of-separation}

\input{conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{mi-estilo-else}
\bibliography{../LSFA/citados}
%\input{separation_arxiv.bbl}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: