### 2019-01-10 LICS/isabelle.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Isabelle}
\label{sec:isabelle}
%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-
\subsection{Logics}
\label{sec:logics}
Isabelle \cite{Isabelle,DBLP:books/sp/Paulson94} provides a
meta-language called \emph{Pure} that consists of a fragment of higher
order logic, where \isatt{\isasymRightarrow} is the function-space
arrow. The meta-Boolean type is called \isatt{prop}. Meta-connectives
\isatt{\isasymLongrightarrow} and \isatt{\&\&\&} fulfill the role of
implication and conjunction, and the meta-binder \isatt{\isasymAnd}
corresponds to universal quantification.

On top of \emph{Pure}, theories/object logics can be defined, with
their own types, connectives and rules. Rules can be written  using
meta-implication: ``$P$, $Q$, and $R$ yield $S$'' can be written
\[
P \ \isatt{\isasymLongrightarrow}\ Q\ \isatt{\isasymLongrightarrow}\ R\ \isatt{\isasymLongrightarrow}\ S
\]
(as usual,  \isatt{\isasymLongrightarrow} associates to the right), and
syntactic sugar is provided to curry the previous rule as follows:
\[
\isasymlbrakk P; Q; R \isasymrbrakk \ \isatt{\isasymLongrightarrow}\ S.
\]
One further example is given by induction on the natural numbers
\isatt{nat},
\[
\isasymlbrakk P(0);\ (\textstyle\isasymAnd
x.\ P(x)\ \isasymLongrightarrow\ P(\isatt{succ}(x))) \isasymrbrakk
\ \isasymLongrightarrow\ P(n), 
\]
where we are omitting the ``typing'' assumtions on $n$ and $x$.

We work in the object theory \emph{Isabelle/ZF}. Two types are defined
in this theory: \tyo, the object-Booleans, and \tyi,
sets. 
%% It must be insisted that the types are defined axiomatically, not
%% recursively. That is, although there are constants and functions that
%% generate elements of both types, neither of them is 
%% \emph{initial}, in the sense that they are not the least types
%% obtained by combining the given constants and operations. 
It must be observed that predicates (functions with arguments of
type \tyi{} with values in \tyo) do not correspond to first-order formulas;
in particular, those are not recursively constructed.
This will
have concrete consequences in our strategy to approach the
development. From the beginning, we had to resort to
\emph{internalized} formulas \cite[Sect.~6]{paulson_2003}, i.e.\ elements of type $\tyi$ that
encode first-order formulas with a binary relation symbol, and the
satisfaction predicate \isatt{sats\,::\,"i\isasymRightarrow i\isasymRightarrow i\isasymRightarrow o"}  between a set
model with an environment and an internalized formula (where the
relation symbol is interpreted as membership). The set 
\isatt{formula::"\tyi"}
 of internalized
formulas is defined by recursion and hence it is possible to perform
inductive arguments using them. In this sense, the object-logic level
is further divided into \emph{internal} and \emph{external}
sublevels.

The source code is written for the 2018 version of Isabelle and can be downloaded
from 
\ifIEEE
\begin{center}
\texttt{https://cs.famaf.unc.edu.ar/\~{}mpagano/forcing/}
\end{center}
\fi
\ifarXiv
\url{https://cs.famaf.unc.edu.ar/~mpagano/forcing/}
\fi
(with
minor modifications, it can be run in Isabelle2016-1). Most of it is
presented in the (nowadays standard) declarative flavour called
\emph{Isar} \cite{DBLP:conf/tphol/Wenzel99}, where intermediate
statements in the course of a proof are explicitly stated,
interspersed with automatic tactics handling more trivial steps. The
goal is that the resulting text, a \emph{proof document}, can be
understood without the need of replaying it and viewing the proof state
at each point.

%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-%-
\subsection{Locales}
\label{sec:locales}
Locales \cite{ballarin2010tutorial} provide a neat facility to
encapsulate a context (fixed objects and assumptions on them) that is
to be used in proving several theorems, as in usual mathematical
practice. Furthermore, locales can be organized in hierarchies.

In this paper, locales have a further use. The \emph{Fundamental
  Theorems of Forcing} we use talk about a specific map $\forceisa$
from formulas to formulas. The definition of $\forceisa$ is involved
and we will not dwell on this now; but applications of those theorems
do not require to know how it is defined. Therefore, we black-box it
and pack everything in a locale called \texttt{forcing\_thms} that
assumes that there is such a 
map that satisfies the Fundamental Theorems.

%% \begin{description}
%% \item[\texttt{forcing\_notion}] preorden con top
%% \item[\texttt{countable\_generic}] lo anterior con una familia contable de densos.
%% \item[\texttt{M\_ZF}] axiomas.
%% \item[\texttt{forcing\_data}]: lo anterior contable transitivo y una notion.
%% \item[\texttt{forcing\_thms}]: eso.
%% \item[\texttt{G\_generic}]: lo anterior y G es genérico.
%% \item[\texttt{M\_extra\_assms}]: check in M e instancia de reemplazo para G.
%% \item[\texttt{G\_generic\_extra}]: los dos anteriores (no sé si sigue estando)
%% \end{description}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: