### 2019-03-07 LICS/rebuttal.txt
limpiando un poco

First and foremost, we would like to thank the referees for their
careful reading and detailed criticism. We are indeed greatly
benefited by their evaluation.

----------------------- REVIEW 1 ---------------------

> Moreover, I would have expected at least a bit of discussion on the
  benefits/drawbacks of choosing Isabelle/ZF over other possible
  frameworks (Coq, Mizar, Isabelle/HOL, HOL Light, ... and the
  corresponding underlying logics).

Insistir que el overhead era tremendo si no usábamos Isabelle/ZF.

> - "the fruitful use of forcing techniques to extend the Curry-Howard
>isomorphism to classical axioms" -> this claim is misleading. The two
>papers you cite show how to express forcing within Krivine
>realizability, a framework which is already classical. Therefore, in
>these papers, forcing techniques are *not* used to extend CH to
>classical logic (which has to do with control operators, not
>forcing). They rather unveil the computational content of forcing
>within Krivine realizability, just like Jaber et al. show how to
>define it within type theory [14,15].  Llamada para el Dr. Pagano...

> - ∈ restricted to M -> this suggests that it is defined outside of
    M, right? Isn't this strange?

Hay que decir que ∈ es el mismo para todos los modelos (el externo).

> - induction on terms of type o is not available -> is this related
    to the chosen framework (Isabelle/ZF) or is it more general?

Aclarar

> - "it is also possible", "it can be shown" -> if you did it, say it
    as such, otherwise I am not sure that this is informative or
    useful (but I might be missing something)

Decir que lo hicimos pero no es parte de esta submission > - Union is
absolute for any transitive class -> is it well-known ?

YES!

----------------------- REVIEW 2 ---------------------

> Summary: This paper presents a formalization in Isabelle of some
> axioms of Set theory of a generic extension of a model of Set theory
> using forcing. The main axiom, according to the authors, that has
> been shown to hold in a generic extension is Separation. The paper
> starts by presenting very briefly Isabelle, axioms of Set theory,
> mentioning the representation of first-order formulas as sets done
> by Paulson and discuss their presentation of renaming of
> formulas. Then the authors explain briefly the generic extension of
> a model of Set theory using forcing, some modifications they have
> done on the development of Paulson of “ZF-Constructible”. The
> authors explain how to prove the “easy axioms” first. Then they
> present the fundamental lemmas of forcing which are given as axioms
> in the formalization. After that, they show that Separation holds in
> the generic extension providing the fundamental lemmas
> holds. Finally, they conclude by discussing how the main missing
> axiom, i.e. Replacement, could be proven.

> Questions to the authors: “We are only able to prove that the
> generic extension satisfies Separation for any particular number of
> parameters, but not in general.”  This looks like a strong
> limitation, can you comment on this.

Hay que decir que no (de hecho con 2 parámetros estamos haciendo
todo), pero tendría que sentarme un rato a confirmar de nuevo esto.

----------------------- REVIEW 3 ---------------------

> My biggest complaint is that this work does not seem to provide the
  kind of fundamental contributions to logic that LICS is looking
  for. There is no new mathematics here, and this isn't even the first
  formalization of separation.

WTF!? Acá hay que preguntar cuál es la otra.

----------------------- REVIEW 4 ---------------------

> The more important point is that the "increment" presented in this
> paper is (IMO) too little for being published in a top conference
> such as LICS. Indeed, the paper only deals with the soundness of
> Separation and Powerset in generic extensions. The formalization of
> the soundness of Extensionality, Regularity, Pairing, Union and
> Infinity has been published before,

Acá la pifió bárbaro. Después repite esto que todos los axiomas
fáciles están en otro paper. ¿Cómo llegó a tener esa impresión?

> The other point deals with the formalization choices of the authors,
> which are never really discussed. It seems to me that the authors
> blindly followed Kunen's presentation of forcing (based on countable
> transitive models and the Rasiowa-Sikorski lemma), without comparing
> it with other presentations that would be (IMO) *much more suitable*
> to a formalization of forcing in a proof-assistant, especially in
> the perspective of deriving relative consistency proofs from this
> work.  (Which is proof-theoretically impossible from Kunen's
> approach, as I explain below in my detailed comments.)

Raro esto. Kunen tiene una sección específica sobre cómo se puede
hacer esto.

> Moreover, the choice of the authors to follow Kunen's approach
> induces many complications in their formalization (for instance: the
> use of internalized formulas) that could have been avoided by taking
> another approach (for instance, the one presented by Jech in his
> textbook on Set theory). Perhaps these complications explain why the
> authors had to split the presentation of their formalization in
> several papers.

> p. 2, second column: internalized formulas. Here you should justify
> why you use internalized formulas, that is: formulas as sets.  IMO,
> it would be more natural to inductively define an Isabelle type "f"
> of first-order formulas (or codes of first-order formulas), together
> with a decoding function that would map each code of a first-order
> formula to a real Isabelle formula. So that the image of this map
> would capture exactly the class you are interested in.  (At least,
> this would be the standard way of doing in Coq or Agda.)

Bueno, se ve que no le quedó claro que esto no se puede hacer. Hay que
insistir.

> Moreover, such a definition would reflect the fact that, in model
> theory, formulas are traditionally considered as external objects:
> they live in the ambient theory, outside all models.

> On the other hand, I agree that "formulas as sets" are useful to
> define the notion of constructible set, and the corresponding
> constructible universe. But the point is that in forcing, we
> (normally) never need such internalized formulas! The theory of
> forcing is independent from the theory of constructible sets, and
> this should appear in a formalization. (Although both theories can
> be nicely combined in particular situations.)

Acá otro pifie, pero es por algo que no aclaramos nosotros: nuestra
presentación no depende de constructibilidad per se, sino de
relativización y absolutez, que es de hecho un prerrequisito para
ambas teorías.

> So why did you consider internalized formulas, and not external
> formulas, inductively defined "outside" as explained above?  I
> understand you want to reuse parts of the (previous) formalization
> of the constructible universe, but this is (scientifically) a very
> weak argument. IMO, considering internalized formulas for defining
> forcing is just introducing unnecessary complication.

Bueno, hacer todo desde cero es una complicación innecesaria para
nosotros (!).

> Moreover, considering internalized formulas (that is: formulas as
> sets) raises another problem, which is that there might be more
> internalized formulas than first-order formulas (such as seen
> outside). Indeed, your axiomatized type "i" of ZF sets is nothing
> but a (hypothetical) model of ZF inside Isabelle. But such a model
> can be non-standard, in the sense that it might contain more natural
> numbers (and thus more formulas-as-sets) than the Isabelle type of
> natural numbers. Note that this problem is harmless in practice, but
> it is worth mentioning it.

Todo está pasando dentro de "i", así que no es relevante esta
observación. Hay que discutirlo un poco más.  > > p. 2, last paragraph
of 2nd column. "Forcing is a technique to extend > very special kind
of models...". Actually, not. There are several > ways to construct
forcing models of ZF, and the one you consider in > your paper (taken
from Kunen's book) is only *one* of them.  > > The method you consider
is based on countable transitive models > (ctms) of ZF, whose main
advantage is that they allow the use of > the Rasiowa-Sikorski lemma
to construct the generic filter.  > Basically, a ctm of ZF captures
three properties: > > (1) M is a set equipped with a binary relation
(to interpret > membership) that is well-founded _from the point of
view of the > outer set-theoretic universe_ (the type "i"). This
property > implies that M fulfills the foundation axiom, but the
converse > is not true. So basically, this property says that the
notion > of well-foundedness is the same inside M and outside M (that
is: > in the outer model "i").  > > (2) M is transitive, which means
that the binary relation interpreting > membership is membership
itself (in the sense of "i"). This point > is actually just a
technical facility, to avoid distinguishing two > membership
predicates.

No sólo eso. Se está olvidando de los resultados de absolutez.  > >
(3) M is countable (from the point of view of the universe "i").  >
This point is crucial to prove the Rasiowa-Sikorski lemma.  > > As you
mention it in Lemma 1, the properties (2) and (3) are >
proof-theoretically free, since once you have (1), you can easily >
enforce (2) and (3) by combining a Mostowski collapse with the >
(descending) Löwenheim-Skolem theorem.  > > But the point is that (1)
is actually very strong, and more > important: it is *not known* to be
equivalent to the consistency > of ZF. By this, I mean that it might
be the case that ZF has > no transitive model while being consistent.

Tranquilate un poco... Esto también es irrelevante, porque sólo
necesitamos un ctm de un fragmento suficientemente grande de ZFC (cosa
que el Teorema de Reflexión nos da).

> This simple argument shows that transitive models of ZF, although
> they are very convenient from the point of view of model theory,
> *cannot* be used to prove relative consistency results about forcing
> (in the proof-theoretic sense), since there is *no known way* to
> turn an arbitrary model of ZF (such as given by the completeness
> theorem of 1st-order logic) into a transitive model.

Ídem... Sólo necesitamos un fragmento finito de esto, como lo explica
Kunen en el libro nuevo.  > > This is why other approaches have been
developed in the literature.  > For instance, the construction of a
Boolean-valued model of names > inside the ground model M, and then a
quotient by a suitable > ultrafilter (see Jech's book for
instance). The main advantage of > this approach is that it can be
applied to _any model_ of ZF. (In > this case, the generic filter is
directly constructed from the > corresponding name, without the
Rasiowa-Sikorski lemma.)  > > However, the ground model M and the
corresponding generic extension > M[G] are in general *not* transitive
models. But from a proof- > theoretic point of view, this property is
definitely useless!  > > So, since you are interested in relative
consistency results (as > mentioned in the introduction of the paper),
why didn't you consider > this alternative approach, that would work
with _any_ model of ZF?  > (Moreover, it seems to me that the job of
formalization would have > been easier.) IMO, the paper is
dramatically missing a discussion > about this important foundational
issue (that has also a dramatic > impact on your formalization
choices).

Me gusta más el enfoque de los ctm para trabajar en general (y
off-the-record, es el que más entiendo de todos), y además tratamos de
seguir lo más posible una sola referencia (i.e. Kunen) para darle
cohesión al trabajo. La discusión sobre la conveniencia de usar el
enfoque de los modelos a valores booleanos es más interesante, pero
creo que otra vez nos pone en el brete de tener que reimplementar
varias cosas.

> Another advantage of constructing the generic extension as a
> quotient of the Boolean-valued model of names (defined as a class
> within M) would be the following. In your work, you need to
> distinguish three models: - the Isabelle type i of sets (the ambient
> model); - the ctm M, that is an object of type i (the ground model);
> - the generic extension M[G].  By considering the approach I
> suggest, you would only need two models: - the type i, which would
> serve as a ground model M as well; - the generic extension M[G],
> that would be defined as a quotient of the class of names (within i
> = M).

> p. 3, second last par. of column 1. "We are only able to prove that
> the generic extension satisfies Separation for any particular number
> of parameters, but not in general." What do you mean precisely?  At
> which level lives the quantification in "any particular number of
> parameters"? A quantification over the set N (inside M and "i")?  Or
> a quantification over the type of natural numbers in Isabelle?
> Please clarify this point.

Aclarar esto.

> p. 9, column 1, 3rd paragraph: "The Replacement Axiom [...] requires
> more work to be done. In Kunen it requires a relativized version
> [...] of the Reflection Principle." IMO, this requirement is another
> drawback of following Kunen's approach. By taking the other method I
> suggest, you would definitely not need the Reflection Principle to
> prove the soundness of Replacement.

Tendría que ver esto, no estoy seguro ya mismo. Este tipo entiende,
sí, pero hay gente que entiende bien un enfoque y no el otro (bueno, a
mí me pasa un poco pero al revés de este muchacho).