### 2019-03-07 LICS/rebuttal.txt
Pasada de Miguel por respuesta
Cambios en las respuestas al 3 y al 4.

First and foremost, we would like to thank the referees for their
careful reading and detailed criticism. We are indeed greatly
benefited by their evaluation.

We first address some issues that appeared in more than one
review. Our choice of Isabelle/ZF as framework was simply based on the
need to decrease the overhead of the formalization process. Indeed,
the main drawback of this choice is that we cannot perform inductive
arguments on "external" formulas (i.e., Isabelle terms of type
o). This is the reason why we are using internalized formulas.

There is are some questions in common on parameters of Separation, but
those are answered below.

Questions by REVIEWER 1
=======================

> - ∈ restricted to M -> this suggests that it is defined outside of
>   M, right? Isn't this strange?

Not really, since ∈ is defined in the "ambient" Isabelle type i, and
one can restrict it to any set.

> - "it is also possible", "it can be shown" -> if you did it, say it
>    as such, otherwise I am not sure that this is informative or
>    useful (but I might be missing something)

We actually did, but that wasn't part of this submission. We are sorry
for the imprecision

> - Union is absolute for any transitive class - is it well-known ?

Yes, indeed.

Questions by REVIEWER 2
=======================

> Questions to the authors: “We are only able to prove that the
> generic extension satisfies Separation for any particular number of
> parameters, but not in general.”  This looks like a strong
> limitation, can you comment on this.

Actually, it is well known [Levy, SchiSchli] that there are
axiomatizations of ZF without any parameters. We have chosen an
axiomatization using at most one parameter in both Separation and
Replacement. In this sense, we can prove that M[G] satisfies the "same
amount of Separation" that M is assumed to satisfy.

Questions by REVIEWER 3
=======================

> Summary:
>
> This paper presents a formalization of the separation axiom from set
> theory using the forcing library that has already been built in
> Isabelle/ZF by Paulson.

This is a mistake. Forcing was not formalized by Paulson.

There are no questions by the third Reviewer, but this Reviewer states
that ours is not the firsgt formalization of separation. This strong
claim is quite misleading: does the Reviewer refer to some
formalization of the Axiom of Separation of some set theory? It's hard
to imagine how does such a formalization look like. We are extremely
eager to know about this, would they provide a reference for such a
strong claim?

Questions by REVIEWER 4
=======================

> The more important point is that the "increment" presented in this
> paper is (IMO) too little for being published in a top conference
> such as LICS. Indeed, the paper only deals with the soundness of
> Separation and Powerset in generic extensions. The formalization of
> the soundness of Extensionality, Regularity, Pairing, Union and
> Infinity has been published before,

There is a mistake in this assertion. The only axiom that has been
published before (at least by us) was Pairing.

> On the other hand, I agree that "formulas as sets" are useful to
> define the notion of constructible set, and the corresponding
> constructible universe. But the point is that in forcing, we
> (normally) never need such internalized formulas! The theory of
> forcing is independent from the theory of constructible sets, and
> this should appear in a formalization.

We agree that internalized formulas are not needed per se; however in
Isabelle/ZF internalized formulas are the only representation of
formulas that are syntactically manipulable. We welcome the suggestion
of extending Isabelle/ZF with a new inductive type "f" for formulas,
this would, however, change the meta-theory (in contrast with Agda, Coq,
or even Isabelle/HOL there is no support in Isabelle/ZF for introducing
inductive types).

Nevertheless, we are *not* using the theory of constructible sets for
our formalization, only the absoluteness results (which are
prerequisites for both theories).

> This simple argument shows that transitive models of ZF, although
> they are very convenient from the point of view of model theory,
> *cannot* be used to prove relative consistency results about forcing
> (in the proof-theoretic sense), since there is *no known way* to
> turn an arbitrary model of ZF (such as given by the completeness
> theorem of 1st-order logic) into a transitive model.

Though it is a fact that Con(ZF) does not imply the existence of ctms
of ZF, finitary consistency proofs can be obtained through the ctm
approach, since one only needs that the ctm M satisfy only a finite
fragment of ZF at a time (and such an M can be obtained by a standard
reflection argument). For a reference, check [Kunen, IV.5].

> p. 3, second last par. of column 1. "We are only able to prove that
> the generic extension satisfies Separation for any particular number
> of parameters, but not in general." What do you mean precisely?  At
> which level lives the quantification in "any particular number of
> parameters"? A quantification over the set N (inside M and "i")?  Or
> a quantification over the type of natural numbers in Isabelle?
> Please clarify this point.

Isabelle/ZF does not define a type of natural numbers (we just have i
and o). "In general" would mean an universal quantification over the
set nat::i.

References
==========

    [Kunen] K. Kunen, Set Theory, 2nd ed., ser. Studies in Logic.
    	    College Publications, 2011, revised edition, 2013.

     [Levy] Azriel Levy, Parameters in the comprehension axiom schemas
     	    of set theory, in Proceedings of the Tarski symposium,
     	    Proceedings of Symposia in Pure Mathematics, vol. 25, American
     	    Mathematical Society, Providence, RI, pp. 309–324.

[SchiSchli] Ralf Schindler and Philipp Schlicht, ZFC without
	    parameters (A note on a question of Kai Wehmeier)
	    http://wwwmath.uni-muenster.de/logik/Personen/rds/ZFC_without_parameters.pdf