### 2019-03-19 ITP/prior-results.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Prior results}

Paulson's work provides an appropriate starting point for our
developments.

There are other formalizations of set theory (Isabelle/HOL, Mizar, Automath).

To the best of our knowledge, only preliminary approaches to the
formalization of forcing exist; v.g. \cite{Quirin}, which has already been
discussed in our previous work \cite{2018arXiv180705174G}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Separation_In_MG"
%%% ispell-local-dictionary: "american"
%%% End: