### 2019-04-01 ITP/Separation_In_MG.tex
typo cerca de nombre Manu

\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Mechanization of Separation in Generic Extensions} %TODO Please add

%\titlerunning{}%optional, please use if title is longer than one line

\author{Emmanuel Gunther}{FAMAF, Universidad Nacional de Córdoba, Argentina}{gunther@famaf.unc.edu.ar}{}{Partially supported by a CONICET scholarship}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional

\author{Miguel Pagano}{FAMAF, Universidad Nacional de Córdoba, Argentina}{pagano@famaf.unc.edu.ar}{}{}%{[orcid]}{[funding]}
\author{Pedro Sánchez Terraf}{CIEM-FAMAF, Universidad Nacional de Córdoba, Argentina}{sterraf@famaf.unc.edu.ar}{}{}%{[orcid]}{[funding]}

\authorrunning{E. Gunther and M. Pagano and P. Sánchez Terraf}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Emmanuel Gunther and Miguel Pagano and Pedro Sánchez Terraf}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{General and reference~General literature}
\ccsdesc[100]{General and reference}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Isabelle/ZF, forcing, names, generic extension, constructibility}%TODO mandatory; please add comma-separated list of keywords

\category{}%optional, e.g. invited paper

\relatedversion{A full version of the paper is available at \url{https://arxiv.org/abs/1901.03313}}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{The whole Isabelle/ZF development is available at \url{https://cs.famaf.unc.edu.ar/\~{}mpagano/forcing/}}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{We would like to thanks to anonymous referees of a
  previous version of this paper that helped us to enhance the
  presentation. We are grateful to Alexandre Miquel for his comments
  on forcing and Boolean-valued models.}%optional

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)
\usepackage{isabelle,isabellesym}
\usepackage{csquotes}
\input{header-draft}
\IEEEtrue

\begin{document}

\maketitle
\input{abstract}

\input{intro}

\input{isabelle}

\input{zfc-axioms}

\input{renaming}

\input{generic-extensions}

\input{recursion}

% \input{prior-results}

\input{hacking}

\input{easy-axioms}

\input{forcing}

\input{proof-of-separation}

\input{conclusion}

\bibliography{../LSFA/citados}

\newpage
\onecolumn
\appendix
\section{A short overview of our development}
\input{appendix}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: