### 2019-12-14 IJCAR2020/conclusion.tex
pedacitos del paper

\section{Conclusion}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: