### 2019-12-14 IJCAR2020/isabelle_metatheory.tex

\section{Isabelle and (meta)theories}

\begin{enumerate}
\item Description of \emph{Pure} (meta-theory).
\item Isabelle/ZF + $\exists M$ ctm of $\ZF(C)$ (theory). 
\item $\formula$ (inner theory).
\item \textbf{Theorem}: $\exists N$ ctm of $\ZF(C)+\neg\CH$.
\end{enumerate}

in comparison to the ``proof theory approach'':

\begin{enumerate}
\item Primitive recursive arithmetic (meta-theory).
\item \textbf{Theorem}: Con($\ZF(C)$) $\implies$ Con($\ZF(C)+\neg\CH$).
\end{enumerate}

and then to the ``type-theoretic approach'':

\begin{enumerate}
\item CiC (meta-theory).
\item CiC (theory, $\geq \ZFC + \exists \kappa$ inaccessible).
\item \textbf{Theorem}: Con($\ZF(C)+\neg\CH$).
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: