### 2019-12-16 IJCAR2020/forcing_in_isabelle_zf.tex
pequeñas ediciones al paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{isabelle,isabellesym}
\input{header-draft}
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
\hypersetup{
  colorlinks,
  urlcolor={blue},
  linkcolor={blue!50!black},
  citecolor={blue!50!black},
}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Formalization of Forcing in Isabelle/ZF%
  \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}%
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Emmanuel Gunther\inst{1} \and
Miguel Pagano\inst{1} \and \\
Pedro Sánchez Terraf\inst{1,2}\orcidID{0000-0003-3928-6942}}
%
\authorrunning{E.~Gunther, M.~Pagano, P.~Sánchez Terraf}
%\authorrunning{E. Gunther et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universidad Nacional de C\'ordoba. 
  \\  Facultad de Matem\'atica, Astronom\'{\i}a,  F\'{\i}sica y
    Computaci\'on. \\
\email{\{gunther,pagano,sterraf\}@famaf.unc.edu.ar}
\and
    Centro de Investigaci\'on y Estudios de Matem\'atica (CIEM-FaMAF),
    Conicet. C\'ordoba. Argentina.
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We formalize the development of forcing in the set theory framework of
Isabelle/ZF. Under the assumption of the existence of countable
transitive model of $\ZF$, we construct a generic extension and show
that the latter also satisfies $\ZF$.
\keywords{Isabelle/ZF \and forcing \and names \and generic extension \and constructibility.}
\end{abstract}
%
%
%
\input{intro}

\input{isabelle_metatheory}

\input{absoluteness}

\input{conclusion}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{forcing_in_isabelle_zf}
\end{document}