### 2019-12-27 IJCAR2020/replacement.tex
nuevas secciones

\section{The axiom of replacement}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: