### 2019-12-28 IJCAR2020/conclusion.tex
ToDo list, typos, nuevos archivos

\section{Conclusion}

TODO:
\begin{enumerate}
\item Implement \emph{Basic Set Theory (BST)} by Kunen in
  Constructible: the use of alternatively Replacement or Powerset to
  prove basic absoluteness and closure resuls.
\item Enhance the automatization of formulas
\item Develop the forcing notions to obtain the independence of $\CH$,
  along with the prerrequisite combinatorial results (v.g.\ the
  $\Delta$-system lemma).
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: