### 2020-01-01 IJCAR2020/proper_extension.tex
redondeando lo de extensión propia

\section{Example of proper extension}

Even when the axioms of $\ZFC$ are proved in the generic extension,
one can't claim that the magic of forcing has taken place unless one
is able to provide some \emph{proper} extension with the \emph{same
ordinals}. After all, one is assuming from starters a model $M$ of $\ZFC$,
and in some trivial cases $M[G]$ might end up to be exactly $M$; this
is where \emph{proper} enters the stage. But, for instance, in the
presence of large cardinals, a model $M'\supsetneq M$ might be an
end-extension of $M$ ---this is were we ask the two models to have the
same ordinals, the same \emph{height}.

Two short theory files contains the relevant
results. \texttt{Ordinals\_In\_MG.thy} shows, using the closure of $M$
under ranks, that $M$ and $M[G]$ share the same ordinals (actually,
ranks of elements of $M[G]$ are bounded by the ranks of their names in
$M$):
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ rank{\isacharunderscore}val{\isacharcolon}\ {\isachardoublequoteopen}rank{\isacharparenleft}val{\isacharparenleft}G{\isacharcomma}x{\isacharparenright}{\isacharparenright}\ {\isasymle}\ rank{\isacharparenleft}x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{lemma}\isamarkupfalse%
\ Ord{\isacharunderscore}MG{\isacharunderscore}iff{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}Ord{\isacharparenleft}{\isasymalpha}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymalpha}\ {\isasymin}\ M\ {\isasymlongleftrightarrow}\ {\isasymalpha}\ {\isasymin}\ M{\isacharbrackleft}G{\isacharbrackright}{\isachardoublequoteclose}
\end{isabelle}

To prove this results, we found it useful to formalize induction over
the relation \isatt{ed}$(x,y) \defi x\in\dom(y)$, which is key
to arguments involving names.
\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ ed{\isacharunderscore}induction{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ {\isasymlbrakk}{\isasymAnd}y{\isachardot}\ \ ed{\isacharparenleft}y{\isacharcomma}x{\isacharparenright}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}y{\isacharparenright}\ {\isasymrbrakk}\ {\isasymLongrightarrow}\ Q{\isacharparenleft}x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}Q{\isacharparenleft}a{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

\texttt{Non\_Constructible.thy} contains the first example of a poset
that interprets the locale
\isatt{forcing{\isacharunderscore}notion}. It is the set of all binary
lists with the partial order given by  ``\isatt{xs} is an extension of
\isatt{ys}.'' The sufficient condition for a proper extension is that
the forcing poset is \emph{separative}: every element has two
incompatible extensions.

\begin{isabelle}
\isacommand{lemma}\isamarkupfalse%
\ incompatible{\isacharunderscore}extensions{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p{\isasymin}list{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isacharparenleft}p\ {\isacharat}\ {\isacharbrackleft}{\isadigit{0}}{\isacharbrackright}{\isacharparenright}\ {\isasymbottom}\ {\isacharparenleft}p\ {\isacharat}\ {\isacharbrackleft}{\isadigit{1}}{\isacharbrackright}{\isacharparenright}{\isachardoublequoteclose}
\end{isabelle}

It is to be noted that both proof-documents comprise mere 99 and 79
(resp.) lines.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: