### 2020-01-04 IJCAR2020/absoluteness.tex
mención de los axiomas

\section{Relativization,  absoluteness, and the axioms}
\label{sec:relat-absol}

The concepts of relativization and absoluteness (due to Gödel, in his
proof of the relative consistency of $\AC$ \cite{godel-L}) 
are both
prerrequisites and powerful tools in working with transitive
models. A \emph{class} is simply a predicate $C(x)$ with at least one
free variable $x$.
The \emph{relativization} $\phi^C(\bx)$ of a set-theoretic
definition
$\phi$ (of a relation such
as ``$x$ is a subset of $y$'' or of a function like $y=\P(x)$) to
a class $C$ is obtained by taking its (first-order) definition and
restricting all of its quantifiers to $C$.

\[
x \sbq^C y \equiv \forall z.\ C(z) \longrightarrow (z\in x
\longrightarrow z\in y)
\]

The new formula $\phi^C(\bx)$ corresponds to what is obtained by defining
the concept ``inside'' $C$. In fact, for a class corresponding to a
set $c$ (i.e.\ $C(x) \defi x \in c$), the relativization $\phi^C$ of a 
sentence $\phi$ is equivalent to the satisfaction of $\phi$ in the
first-order model $\lb c, \in\rb$.

It turns out that many concepts mean the
same after relativization to a nonempty transitive class $C$; formally
\[
\forall\bx.\ C(\bx) \longrightarrow (\phi^C(\bx) \longleftrightarrow
\phi(\bx))
\]
When this is the case, we say that the relation defined by $\phi$ is
\emph{absolute for transitive models}. (The absoluteness of functions
moreover requires that
the relativized definition also behaves functionally over $C$.) As
examples, the relation of inclusion $\subseteq$ ---and actually, any
relation defined by a formula (equivalent to one) using only bounded
quantifiers 
$(\forall x\in y)$ and $(\exists x\in y)$--- is absolute for
transitive models. On the contrary, this is not the case with the powerset
operation.

A benefit of the work with transitive models is that [one concept of
  ordinal accross the universe / $M$ / $M[G]$]

A part of this project is to refactorize Paulson's formalization
\cite{paulson_2003} of Gödel's \cite{godel-L}. The main objective is
to maximize applicability of the relativization machinery by adjusting
the hypothesis of a greater part of it early development. Paulson's
architecture had only in mind the consistency of $\ZFC$, but, for
instance, in order to apply it in the development of forcing, too much
is assumed at the beginning; more seriously, some assumptions can't be
regarded as ``first-order'' (v.g. the Replacement Scheme).

The version of \isatt{ZF-Constructible} we present weakens the
assumptions of [many / a dozen / several] absoluteness theorems to that
of a nonempty transitive class.
%
\begin{framed}
  New absoluteness results on powersets --- for \isatt{Least} and
  \isatt{Vset} see Section~\ref{sec:axioms-replacement-choice}
\end{framed}

Apart from the axiom schemes, the rest of the $\ZFC$ 
axioms are stated as predicates on classes (that is,
of type \isatt{(i{\isasymRightarrow}o){\isasymRightarrow}o}); some
of them were already defined by Paulson, and this formulation allows a
better interaction with \isatt{ZF-Constructible}. 
The axioms of Pairing, Union, Foundation,
Extensionality, and Infinity are relativizations of the respective
traditional 
first-order sentences to the class argument. For the Axiom of Choice
we selected a version best suited for the work with transitive
models: the relativization of a sentence stating that for every $x$
there is surjection from an ordinal onto $x$. Finally, Separation and
Replacement were treated separately to effectively obtain first-order
versions. It is to be noted that predicates in Isabelle/ZF are akin to
second order variables and thus do not correspond to first-order
formulas; in particular, there is no induction principle for functions
of type \isatt{i{\isasymRightarrow}o}. For that reason, Separation and
Replacement predicate \emph{on the satisfaction} of a formula $\phi$.
With respect to our previous \cite{2019arXiv190103313G}, we lifted the
arity restriction for the parameter $\phi$ on these schemes,
streamlining various proofs as an immediate benefit and avoiding the
need for tupling.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: