### 2020-01-25 IJCAR2020/conclusion.tex

\section{Conclusion and future work}
\label{sec:conclusion}

We consider that the formalization of the definition of $\forceisa$
and its recursive characterization of forcing for atomic formulas is
a turning point
in our project; the reason for this is that all further
developments will not involve such a daunting metamathematical
component. Even the proofs of the Fundamental Theorems of Forcing
turned out to follow rather smoothly after this initial setup was
ready, the only complicated affair being to show that various dense sets belong
to $M$. 
%
Actually, this is a point to be taken care of: For every new
concept that is introduced, some lemmas concerning 
relativization and closure must be proved to be able to synthesize its
internal definition. Further automation must be developed for this
purpose.

In the course of obtaining internal formulas for the atomic case of
forcing, a fruitful discussion
concerning complementary perspectives on the role of proof assistants
took place. An earlier approach relied more heavily in formula
synthesis, thus making the Simplifier an indispensable main
character. Following this line was quickier from the coding point of
view since few new primitives were introduced and thus fewer lemmas
concerning absoluteness and arities. On the downside, processing was a
bit slower, the formulas synthesized were gigantic and the process on
a whole as more error-prone. In fact, this approach was unsuccessful
and we opted for a more detailed engineering, defining all
intermediate steps. So the load on the assistant, in this part of the
development, balanced from code-production to code-verification.

The next task in our path is pretty clear: To develop the forcing
notions to obtain the independence of $\CH$ 
along with the prerequisite combinatorial results, v.g.\ the
$\Delta$-system lemma. A development of cofinality is under way in a
joint work with E.~Pacheco Rodríguez, which is needed for a general
statement of the latter.

In a second release of \isatt{ZF-Constructible-Trans}, we intend to conform
it to the lines of \emph{Basic Set Theory (BST)}
\cite[I.3.1]{kunen2011set} in which elementary results have proofs
using alternatively Powerset or Replacement. The interest in this
arises because many natural set models (rank-initial segments of the
universe or the family $H(\kappa)$ of sets  of cardinality
less than $\kappa$ hereditarily) satisfy one of those axioms and not the
other. There are also still some older or less significant proofs
written in tactical (\textbf{apply}) format; we hope we will find the
time to translate them to Isar. Finally, the automation of formula
synthesis is on an early stage of development; finishing that module
will make writing our proofs of closure under various operations
faster, and also turn the set theory libraries  more
usable to other researchers.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "forcing_in_isabelle_zf"
%%% ispell-local-dictionary: "american"
%%% End: