### 2020-03-09 IJCAR2020/rebuttal.txt
primera version de rebuttal

We would like to thank the referees for their
careful reading and detailed criticism.

> ----------------------- REVIEW 1 ---------------------
> 
> The paper's main limitation is its treatment of related work. On
> general principles, the main discussion of related work should be near
> the end of the paper, where it could be compared with the work just
> presented. There, it would be appropriate to contrast what was done in
> Paulson's work on constructibility with the current work on
> forcing. More to the point, the recent work by Han and van Doorn on
> forcing in Lean deserves more discussion.

We will gladly incorporate this changes into a revised version.

> ----------------------- REVIEW 2 ---------------------
> 
> In a single
> publication, Han and van Doorn using forcing to show the independence
> of the continuum hypothesis, whereas the authors of this paper are on
> their third publication, and still seem to be at least one paper away
> from showing any independence result.

We believe that the previous assertions are not entirely accurate. As
it is clear from our introduction, our second manuscript on the
subject is not published (and it won't be) in any venue other than the
arXiv. (As a side note, Han & van Doorn have two papers detailing
the full account of their formalization.)

> The main result in this paper is
> simply that forcing extensions satisfy replacement and choice, which
> seem to be fairly straightforward.

In some sense, this is a misstatement of our achievements. The "Main
theorem" of our paper is just the icing on the cake, since there is no
other way to prove this result other than formalizing the machinery of
forcing, which is the real achievement of this paper. Han & van Doorn
[CPP 2020] hint that the formalization of forcing in set theoretical
foundations is practically impossible.

> 
> 
> ----------------------- REVIEW 3 ---------------------
> This work presents a formalization of forcing in Isabelle/ZF and the 
> proof of a first lemma about the forcing relation.

We would like to point out that we proved the fundamental theorems of
forcing, which are the main tools to work with the forcing relation.