### 2020-03-12 IJCAR2020/rebuttal.txt
version final de rebuttal

We would like to thank the referees for their
careful reading and detailed criticism.

> ----------------------- REVIEW 1 ---------------------
> 
> The paper's main limitation is its treatment of related work. On
> general principles, the main discussion of related work should be near
> the end of the paper, where it could be compared with the work just
> presented. There, it would be appropriate to contrast what was done in
> Paulson's work on constructibility with the current work on
> forcing. More to the point, the recent work by Han and van Doorn on
> forcing in Lean deserves more discussion.

We will gladly incorporate these changes and discussion on the
questions posed into a revised version.

> ----------------------- REVIEW 2 ---------------------
> 
> In a single
> publication, Han and van Doorn using forcing to show the independence
> of the continuum hypothesis, whereas the authors of this paper are on
> their third publication, and still seem to be at least one paper away
> from showing any independence result. In particular, CH will require
> some careful cardinal combinatorics.

We knew of Han and van Doorn when we were somewhat advanced in our
project and decided to continue because of the differences:

 1) a weaker meta-theory (Han & van Doorn [CPP 2020] hint that the
 formalization of forcing in set theoretical foundations is practically
 impossible); and

 2) an approach to forcing (ctm) used in a important fraction of the
 literature.

We agree that the remaining work towards the independence of CH will
require one more paper. Notice that, in fact, Han & van Doorn have two
papers (ITP2019, CPP2020) detailing the full account of their
formalization. (As a side note, our second manuscript on the subject
is not published.)

> The main result in this paper is
> simply that forcing extensions satisfy replacement and choice, which
> seem to be fairly straightforward.

This is a misstatement of our achievements. Our "Main Theorem" may
seem straightforward thanks to the more substantial contribution of
the paper: the definition of the forcing relation and the proof of the
forcing theorems. The Main Theorem is a succinct way to demonstrate
the forcing machinery, since there is no other way to produce
extensions of ctms.

> ----------------------- REVIEW 3 ---------------------
> This work presents a formalization of forcing in Isabelle/ZF and the 
> proof of a first lemma about the forcing relation.

We would like to point out that we proved the fundamental theorems of
forcing, which are the main tools to work with the forcing relation.