### 2020-03-13 IJCAR2020/rebuttal-miguel.txt

We would like to thank the referees for their
careful reading and detailed criticism.

> ----------------------- REVIEW 1 ---------------------
>
> The paper's main limitation is its treatment of related work. On
> general principles, the main discussion of related work should be near
> the end of the paper, where it could be compared with the work just
> presented. There, it would be appropriate to contrast what was done in
> Paulson's work on constructibility with the current work on
> forcing. More to the point, the recent work by Han and van Doorn on
> forcing in Lean deserves more discussion.

We will gladly incorporate this changes into a revised version.

> ----------------------- REVIEW 2 ---------------------
>
> In a single publication, Han and van Doorn using forcing to show the
> independence of the continuum hypothesis, whereas the authors of
> this paper are on their third publication, and still seem to be at
> least one paper away from showing any independence result.

We knew of Han and van Doorn when we were somewhat advanced in our
project and decided to continue because of the differences: a
different meta-theory and a different approach to forcing. We think
that this paper is interesting per se (we can agree that the arXiv
paper is a bit premature) as it establish a ground on which to do
forcing arguments.

> The main result in this paper is simply that forcing extensions
> satisfy replacement and choice, which seem to be fairly
> straightforward.
> ----------------------- REVIEW 3 ---------------------
> This work presents a formalization of forcing in Isabelle/ZF and the
> proof of a first lemma about the forcing relation.

We would like to point that the main result may seem straightforward
thanks to the more substantial contribution of the paper: the
definition of the forces function and the proof of the forcing
theorems. Han & van Doorn [CPP 2020] hint that the formalization of
forcing using the ctm and in a set theoretical foundations is more
intricate than via the Boolean-valued models on type-theory.