### 2022-03-26 JSL/independence_ch_isabelle.tex
Random draft for JSL paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{isabelle_indepCH,isabellesym_indepCH}
\input{header-indepCH}
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
\hypersetup{
  colorlinks,
  urlcolor={blue},
  linkcolor={blue!50!black},
  citecolor={blue!50!black},
}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Some lessons after the formalization of the ctm approach to forcing%
  \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}%
}
%
\titlerunning{Lessons after formalizing ctm forcing}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Emmanuel Gunther\inst{1} \and
Miguel Pagano\inst{1} \and \\
Pedro Sánchez Terraf\inst{1,2}%\orcidID{0000-0003-3928-6942}
\and
Matías Steinberg\inst{1}
}
%
\authorrunning{E.~Gunther, M.~Pagano, P.~Sánchez Terraf, M.~Steinberg}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universidad Nacional de C\'ordoba. 
  \\  Facultad de Matem\'atica, Astronom\'{\i}a,  F\'{\i}sica y
  Computaci\'on.
  \and
    Centro de Investigaci\'on y Estudios de Matem\'atica (CIEM-FaMAF),
    Conicet. C\'ordoba. Argentina. \\
    \email{\{gunther,pagano,sterraf\}@famaf.unc.edu.ar\\
        matias.steinberg@mi.unc.edu.ar}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  We'll discuss some highlights of our computer-verified
  proof of the construction, given a countable transitive set model $M$
  of $\ZFC$, of generic extensions  satisfying $\ZFC+\neg\CH$ and $\ZFC+\CH$.
  In particular,
  we isolated a set $\Delta$ of 39 instances
  of Axiom of Replacement and a function $F$
  such that such that for any finite fragment $\Phi\sbq\ZFC$,
  $F(\Phi)\sbq\ZFC$ is also finite and if
  $M\models F(\Phi) + \Delta$ then $M[G]\models \Phi + \neg \CH$.
  We also obtained the formulas yielded by the Forcing Definability Theorem
  explicitly.

  To achieve this, we worked in the proof assistant \emph{Isabelle},
  basing our development on the theory Isabelle/ZF by L.~Paulson and
  others.

  %% The vantage point of the talk will be that of a mathematician but
  %% elements from the computer science perspective will be
  %% present. Perhaps some myths regarding what can effectively be done
  %% using  proof assistants/checkers will
  %% be dispelled.

  We'll also compare our formalization with the recent one by Jesse
  M.~Han and Floris van Doorn in the proof assistant \emph{Lean}.

  \keywords{forcing \and Isabelle/ZF \and countable transitive models
    \and absoluteness \and generic extension \and constructibility.}
\end{abstract}
%
%
%
\input{intro}
%% 
%% \input{isabelle_metatheory}
%% 
%% \input{absoluteness}
%% 
%% \input{forces_definition}
%% 
%% \input{forcing_theorems}
%% 
%% \input{proper_extension}
%% 
%% \input{replacement_choice}
%% 
\input{random}

\input{related}

\input{lessons}
%%
%% \input{conclusion}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
%\bibliographystyle{splncs04}
\bibliographystyle{mi-estilo-else}
\bibliography{independence_ch_isabelle,clam2021}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: