### 2022-03-26 JSL/random.tex

\subsection{Random comments}

\begin{enumerate}
\item Satisfaction of having one's material fully formalized
  (v.g. cofinality).
\item The concept of absoluteness (\verb|_abs|) splitted into two components
  during the formalization: \verb|is_foo_iff| and \verb|foo_rel_abs| (this last one
  being the concept of absoluteness from  “mathematical practice”) and
  the first one involves a purely relational presentation, introduced
  by Paulson.
\item
  In order to extract information concerning the proof, the need for a
  proof assistant with a more computational flavour (Agda, Coq)
  arises! As a second thought, obviously.
\item  \textbf{Pros and cons of an untyped environment}: Closer to the
  set-theoretic point of view/prone to the most stupid errors.
\item \textbf{“Pros” and cons of lack of automation}: we had to be extremely
  detailed, which provides explicit information / the formalization
  work was pain in the neck and much lower.
\item In order to extract information concerning the proof, the need for a
  proof assistant with a more computational flavour (Agda, Coq)
  arises! As a second thought, obviously.
\item Promotion of formalization by Buzzard.
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: