### 2022-04-27 JSL/proof_assists_isabelle.tex

\section{Proof assistants and Isabelle/ZF}
\label{sec:proof-assist-isabelle}

Proof assistants provide diverse aids for the task
of formalizing a piece of mathematics. They are usually implemented
using a typed programming language; rigor is enforced by defining a
type of ``theorems'', whose members can only be constructed using
operations stipulated in a small \emph{kernel} which encodes the
underlying foundation of the assistant. Further developments interact
with the type of theorems only through the kernel, and thus the latter
is the only ``trusted'' part of the assistant's code.

Several of the more established assistants (HOL Light, Coq, Isabelle)
are programmed in some variant of the ML language (which was
originally designed for this purpose); the newer Lean, on the other
hand, was originally conceived as a standalone functional programming
language with all the features of a standard assistant.

\subsection{Four layers}

In the case of Isabelle, Standard ML is the deeper layer of the
assistant, in which the kernel is written.

%% Isabelle \cite{Isabelle,DBLP:books/sp/Paulson94} is a general proof
%% assistant based on fragment of higher-order logic called
%% \emph{Pure}. 
%% The results presented in this work are theorems of a
%% version of $\ZF$ set theory (without the Axiom of Choice, $\AC$) 
%% called \emph{Isabelle/ZF}, which is one of the
%% ``object logics'' that can be defined on top of Pure (which is then
%% used as a language to define rules). Isabelle/ZF defines types
%% \isatt{i} and \isatt{o} for sets and Booleans, resp., and the $\ZF$
%% axioms are written down as terms of type \isatt{o}.
%% 
%% It should be noted that Pure is a very weak framework and has no
%% induction/recursion capabilities. So the only way to define functions
%% by recursion is inside the object logic. (This works the same for
%% Isabelle/HOL.) For this reason, to define the relation of forcing, we
%% needed to resort to \emph{internalized} first-order formulas: they
%% form a recursively defined set \isatt{formula}. For example, the
%% predicate of satisfaction
%% \isatt{sats::i{\isasymRightarrow}i{\isasymRightarrow}i{\isasymRightarrow}o}
%% (written $M,\mathit{ms}\models\phi$ for a set $M$,
%% $\mathit{ms}\in\isatt{list}(M)$ and $\phi\in\formula$)
%% had already been defined by recursion in Paulson~\cite{paulson_2003}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: