### 2022-05-04 JSL/proof_assists_isabelle.tex
Isabelle/ZF, Constructible, and internalized formulas

\section{Proof assistants and Isabelle/ZF}
\label{sec:proof-assist-isabelle}

Proof assistants provide diverse aids for the task
of formalizing a piece of mathematics. They are usually implemented
using a typed programming language; rigor is enforced by defining a
type of ``theorems'', whose members can only be constructed using
operations stipulated in a small \emph{kernel} which encodes the
underlying foundation of the assistant. Further developments interact
with the type of theorems only through the kernel, and thus the latter
is the only ``trusted'' part of the assistant's code.

Several of the more established assistants (HOL Light, Coq, Isabelle)
are programmed in some variant of the ML language (which was
designed for this purpose); the newer Lean, on the other
hand, was originally conceived as a standalone functional programming
language with all the features of a standard assistant.

In the case of Isabelle, Standard ML is the first of the four layers
in which we work in this assistant. Both the kernel and automation
of proofs is coded in ML, sometimes as a substitute for induction on
formulas, as the next section explains.

\subsection{Isabelle metalogic $\Meta$}

The second layer of Isabelle is an
intuitionistic fragment of higher-order logic (or simple type theory)
called $\Meta$; its original version was described in \cite{Paulson1989},
and the addition of “sorts” appears in \cite{Nipkow-LF-91}.

The only predefined type is $\prop$ (“propositions”); new base types
can be postulated when defining objects logics. Types of higher order can be
assembled using the function space constructor $\fun$.

The type of propositions $\prop$ equipped with a binary operation
$\implies$ (“meta-implication”) that is used to represent the object
logic rules. As an example, the axiomatization of first-order logic
postulates a type $\tyo$ of booleans, and Modus Ponens
% https://isabelle.in.tum.de/dist/library/FOL/FOL/IFOL.html#IFOL.mp|axiom
is written as
\[
  [P\limp Q] \implies ([P] \implies [Q]).
\]
The square brackets (which are omitted in Isabelle theories) represent
an injection from $\tyo$ into $\prop$. % ($[P]$ can be read as “$P$
% holds”)
A consequence of this representation is that every formula of
the object logic appears as atomic for $\Meta$.

%% Quantification is handled in the meta-level using a functional $\ALL$
%% with polymorphic type $(\alpha \fun \prop) \fun \prop$. 
Types in Isabelle are organized into \emph{classes} and \emph{sorts};
for ease of exposition, we will omit the former.  The axiomatization
of first-order logic postulates a sort $\{\type{term}\}$ (of
“individuals,” or elements of a first-order universe of discourse) and
stipulates that every further type variable $\alpha$ must be of that
sort. In particular, Isabelle/ZF only postulates one new type $\tyi$
(“sets”) of sort $\{\type{term}\}$. Hence, from the type of the universal
quantifier functional $\forall :: (\alpha \fun \tyo) \fun \tyo$ it
follows that it may only be applied to predicates with a variable of
type $\tyi$. This ensures that the object logic is effectively
first-order.

Paulson  \cite{Paulson1989} carried out a proof that the encoding
$\Meta_{\mathrm{IFOL}}$ of
intuitionistic first-order logic IFOL without equality  in the original $\Meta$ is
conservative (there is a correspondence between provable $\phi$ in
IFOL and provable $[\phi]$ in $\Meta_{\mathrm{IFOL}}$) by putting
$\Meta_{\mathrm{IFOL}}$ proofs in \emph{expanded normal form}
\cite{MR0387024}. Passing to classical logic does not present
difficulties, but the treatment of meta-equality differs between the
original and the present incarnation of $\Meta$; details for the
latter are exhaustively expounded in the recent formalization by
Nipkow and Roßkof \cite{10.1007/978-3-030-79876-5_6}.

The meta-logic $\Meta$ is rather weak; it has no induction/recursion
principles. Types are not inductively presented and, in particular, it
is not possible to perform proofs by induction statements about
object-logic formulas (which are construed as terms of type $\tyi \fun
\dots \fun \tyi \fun \tyo$). Two ways to overcome this limitation are:
\begin{enumerate}
\item
  to construct the
  proof of each instance of the statement by hand or by programming on
  ML; or 
\item
  to encode formulas as sets and prove an internal version statement
  using induction of $\ZF$.
\end{enumerate}

For recursive definitions, only the second option is available, and
that is the way the Definition of Forcing is implemented in our
formalization.

\subsection{Isabelle/ZF}

For the most part, the development of set theory in Isabelle is
carried out using its ZF object logic
\cite{DBLP:journals/jar/PaulsonG96}. Apart from the type and sort
declarations detailed above, it features a finite axiomatization,
% https://isabelle.in.tum.de/dist/library/FOL/FOL/IFOL.html#ZF_Base.mem|const
with a predicate for membership, constants for the empty set and an
infinite set, and functions $\isatt{Pow}::\tyi\fun\tyi$,
$\union::\tyi\fun\tyi$, and $\isatt{PrimReplace} :: \tyi \fun (\tyi
\fun \tyi \fun\tyo)\fun \tyi$ (for Replacement). The Axiom of
Replacement
% https://isabelle.in.tum.de/dist/library/FOL/FOL/IFOL.html#ZF_Base.replacement|axiom
has a free predicate variable $P$: % $P ::\tyi\fun\tyi\fun\tyo$:
\begin{multline*}
  (\forall x \in A .\ \forall y\, z.\ P(x, y) \wedge P(x, z)
  \longrightarrow y=z) \implies \\
  b \in \isatt{PrimReplace}(A, P)
  \longleftrightarrow(\exists x \in A .\ P(x, b)) 
\end{multline*}
The restrictions on sorts described above ensure that it is not
possible that higher-order quantification gets used in $P$. The
statement of $\AC$ also has a higher-order free variable, but we only
use $\AC$ for demonstration purposes.

Isabelle/ZF reaches Hessenberg's $|A|\cdot|A| = |A|$. Our decision
(during 2017) to
use this assistant was triggered by its constructibility
library \citep{paulson_2003},
% https://isabelle.in.tum.de/dist/library/ZF/ZF-Constructible/
which contains the development of $L$, the proof that it satisfies
$\AC$, and a version of the Reflection Principle. The latter was
actually encoded as a series of instructions to Isabelle automatic
proof tools that would prove each particular instance of reflection:
This is an example of what was said at the end of the previous
section.

The development of relativization and absoluteness for classes $C::
\isatt{i} \fun \isatt{o}$ follows the same pattern. Each particular
concept was manually written in a relational form and relativized.
Here, the contrast between the usual way one regards $\ZF$ as
first-order theory in the language $\{\in \}$ and the mathematical
practice of freely using defined concepts comes to the
forefront. Assistants have refined mechanisms to cope with defined
concepts (which also make their foundations a more complicated
than plain first-order logic), and this is the only that  nontrivial
mathematics can be formalized. But when studying relative interpretations, one
usually assumes a spartan syntax and defines relativization by
induction of formulas of the more succinct language. The approach
taken in \session{ZF-Constructible} is to consider relativizations of
the formulas (of type $\tyo$) that define each concept. For instance,
in the case of unions, we find a relativization
$\isatt{big{\uscore}union}:: (\isatt{i} \fun \isatt{o}) \fun \isatt{i}
\fun \isatt{i} \fun \isatt{o}$ of the statement
“$\union A = z$”:
\[
 \isatt{big{\uscore}union}(M,A,Z) \equiv \forall x[M].\ x \in z
 \longleftrightarrow (\exists y[M].\ y \in A \land x \in y)
\]
where $\forall x[M]\dots$ stands for $\forall x.\ M(x)\limp \dots$,
etc. The need to work with \emph{relational} presentations of defined
concepts stems from the fact that the model-theoretic definition of
$L$ requires working with set models and satisfaction, which is
defined for codes of formulas.

\subsection{Internalized formulas}

\session{ZF-Constructible} defines the set $\formula$ of codes for
first-order formulas. These, alongside with lists, are instances of
Isabelle/ZF treatment of inductively defined datatypes; induction
and recursion theorems for them are proved automatically.

To avoid problems related to the binding of variables, de Bruijn
indices are used instead. The satisfaction predicate
$\isatt{sats}::\tyi\fun\tyi\fun\tyi\fun\tyo$ takes as arguments a set $M$, a list
$\mathit{ms}\in\isatt{list}(M)$ for the assignment of free indices,
and $\phi\in\formula$, and it is written
$M,\mathit{ms}\models\phi$ in our formalization. This completes the
description of the fourth and last formal layer of the development.

Internalized formulas for most (but not all) of the relational
concepts is obtained by guiding the automatic tactics. Actually,
great many of the concepts in \session{ZF-Constructible} where
internalized by hand, as is the case for union,
%% \begin{multline*}
%%   \isatt{big{\uscore}union{\uscore}fm}(A,z) \equiv \\
%%   ((\cdot\forall\cdot\cdot0 \in succ(z)\cdot \longleftrightarrow
%%   (\cdot\exists\cdot\cdot0 \in succ(succ(A))\cdot \land \cdot1 \in
%%   0\cdot\cdot\cdot)\cdot\cdot)
%% \end{multline*}
for which we have the following satisfaction lemma:
\begin{multline*}
  x \in \omega \implies y \in \omega \implies \mathit{env} \in \isatt{list}(A)
  \implies \\
  (A, \mathit{env} \models \isatt{big{\uscore}union{\uscore}fm}(x,y)%\cdot\union x \isatt{ is } y\cdot
  ) \longleftrightarrow  \isatt{big{\uscore}union}(\isatt{\#\#} A, \isatt{nth}(x,
  \mathit{env}), \isatt{nth}(y, \mathit{env}))
\end{multline*}
Note that $x$ and $y$ above are de Bruijn indices,
$\isatt{nth}(x,\mathit{env})$ is the $x$th element of $\mathit{env}$
and $\isatt{\#\#}A::\tyi\fun \tyo$ is the class corresponding to the
set $A::\tyi$.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: