### 2022-05-08 JSL/random.tex
Minor edits

\section{Random comments}

\begin{enumerate}
\item  \textbf{Pros and cons of an untyped environment}: Closer to the
  set-theoretic point of view/prone to the most stupid errors.
\item \textbf{“Pros” and cons of lack of automation}: we had to be extremely
  detailed, which provides explicit information / the formalization
  work was pain in the neck and much lower.
\item In order to extract information concerning the proof, the need for a
  proof assistant with a more computational flavour (Agda, Coq)
  arises! As a second thought, obviously.
\end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: