### 2022-05-27 Independence_CH/conclusion.tex
More lessons into slightly-sensical paragraphs

\section{Conclusion}
\label{sec:conclusion}

One of our main goals was to provide a \emph{proof document}, which
can hopefully be read by mathematicians/set-theorists.

Where did we fail on this?
\begin{enumerate}
\item
  The definition of forces is too codish.
\item
  Proving instances of axiom schemes.
\end{enumerate}

The experience of formalizing math:
\begin{enumerate}
\item It is fun, almost addictive.
\item Feeling of accomplishment after seeing your writings
  validated beyond reasonable doubt (v.g. cofinality).
\item On the grim side: It is easy to forget to put perspective on a
  formalization issue and meditate on paper.
\end{enumerate}

\section*{Acknowledgments}
\label{sec:acknowledgments}
We thanks Z.~Vidnyánszky for some discussions where the idea of
building the summary in \theory{Definitions\_Main} took shape.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: