### 2022-09-12 Independence_CH/discipline.tex
Bye isatt, bye slanted

\section{Discipline for relativization}
\label{sec:discipline-relativization}

As we said in Sec.~\ref{sec:relat-vers-non-absol}, in order to force
$\CH$ and its negation we depended on having relativized versions of
cardinals, Alephs, etc. It was clear for us that our efforts would be
more efficient if we set up a discipline for relativizing sets (terms
of type $\tyi$) and predicates/relations (terms of type $\tyo$).

Paulson only had, for each set, the relational version. It seemed
clearer to us to have a functional version of the relativized concept.
Going back to our example in \ref{sec:tools-relativization}, for the
concept $\isa{cardinal}::\tyi \fun \tyi$ we want its relative
version
$\isa{cardinal{\uscore}rel}::(\tyi \fun \tyo) \fun \tyi \fun\tyi$
and the relational version of the latter
$\isa{is{\uscore}cardinal}::(\tyi \fun \tyo) \fun \tyi \fun \tyi
\fun \tyo$.

Our first attempt of defining a discipline was inspired by
mathematical considerations: if we might prove that
$\isa{is{\uscore}cardinal}$ is functional and also prove the
existence of a witness $\isa{c}$ such that $\isa{M(c)}$ and
$\isa{is{\uscore}cardinal(M,x,c)}$ then
$\isa{cardinal{\uscore}rel(M,x)}$ can be obtained by the operator of
definite descriptions.

Soon we realized that resorting to definite descriptions was needed
only for the most primitive concepts. In fact, once we have a
relativized concept, we can use it to define other relativizations.
For instance, $\isa{cardinal{\uscore}rel}$ depends on having
relative versions of $\isa{bij}$. Instead of relationalizing
$\isa{bij}$ to get $\isa{is{\uscore}bij}$ and then prove
uniqueness and existence of a witness, we define
$\isa{bij{\uscore}rel}$ using $\isa{inj{\uscore}rel}$ and
$\isa{surj{\uscore}rel}$.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: