### 2022-09-20 Independence_CH/replacement_instances.tex

\section{22 replacement instances to rule them all}
\label{sec:repl-instances-appendix}

In Table~\ref{tab:instances1} we show the name of the fourteen
formulas involved in the twenty two instances of replacement needed in
our mechanization. The formulas marked with ($\dagger$) are needed
twice: one by themselves and the other as the argument for
$\isa{ground{\uscore}repl{\uscore}fm}$. The
$\isa{ground{\uscore}repl{\uscore}fm}$ function maps $\phi$ to $\psi$
as described in Sect.~\ref{sec:repl-instances}. These eight instances form
the set \isa{instances3{\uscore}fms}.

% These are expressed using the following predicate:
% \[
%   \isa{ground{\uscore}replacement{\uscore}assm}(M,\mathit{env},\phi)
% \]
% defined as
% $\isa{replacement{\uscore}assm}(M,\mathit{env},\isa{ground{\uscore}repl{\uscore}fm}(\phi))$.
\newcommand{\groundRepl}{\ensuremath{{}^\dagger}}
\newcommand{\replInstSet}[1]{\multicolumn{3}{@{}l}{\hspace{0.3em}{$\mathbf{\mathit{#1}}$}}}

\begin{table}[!h]
\centering
\begin{threeparttable}
\begin{tabular}{r<{\stepcounter{replInstCount}\thereplInstCount.} >{\hspace{2pt}}l @{\hspace{1ex}} p{6cm}}
  \toprule
  \multicolumn{1}{r}{No.} & Formula's name & Comment \\
  \midrule
  \replInstSet{instances1{\uscore}fms}\\
%  \addlinespace
  & \isa{eclose{\uscore}closed{\uscore}fm} \groundRepl & Closure under iteration of $X\mapsto\union X$. \\
  & \isa{eclose{\uscore}abs{\uscore}fm} \groundRepl & Absoluteness of the previous construction.\\
%  \addlinespace
  & \isa{wfrec{\uscore}rank{\uscore}fm} \groundRepl &  For $\in$-rank.\\
%  \addlinespace
  & \isa{transrec{\uscore}VFrom{\uscore}fm} \groundRepl & For rank initial segments.\\
  \midrule
  \replInstSet{instances2{\uscore}fms}\\
%  \addlinespace
  & \isa{wfrec{\uscore}ordertype{\uscore}fm} \groundRepl & Well-founded recursion for the construction of ordertypes. \\
  & \isa{omap{\uscore}replacement{\uscore}fm} \groundRepl & Auxiliary instance for the definition of ordertypes. \\
%  \addlinespace
  & \isa{ordtype{\uscore}replacement{\uscore}fm} \groundRepl& Replacement through the ordertype function, for Hartogs' Theorem.\\
%  \addlinespace
  & \isa{wfrec{\uscore}Aleph{\uscore}fm} \groundRepl& Well-founded recursion to define Aleph.\\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}fms}\\
%  \addlinespace
  & \isa{wfrec{\uscore}Hcheck{\uscore}fm} & Well-founded recursion to define check.\\
  & \isa{wfrec{\uscore}Hfrc{\uscore}at{\uscore}fm}. & Well-founded recursion for the definition of forcing for atomic formulas.\\
  & \isa{lam{\uscore}replacement{\uscore}check{\uscore}fm} & Replacement through $x\mapsto \lb x,\check{x}\rb$, for $\punto{G}$.\\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}notCH{\uscore}fms}\\
%  \addlinespace
  &   \isa{recursive{\uscore}construction{\uscore}fm} &
  Recursive construction of sets using a choice function.\\
  & \isa{recursive{\uscore}construction{\uscore}abs{\uscore}fm} &
  Absoluteness of the previous construction.\\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}CH{\uscore}fms}\\
%  \addlinespace
  & \isa{dc{\uscore}abs{\uscore}fm} &  Absoluteness of the recursive construction in the proof of
  Dependent Choice from $\AC$. \\
  \midrule
  \replInstSet{instances3{\uscore}fms}\\
%  \addlinespace
  \multicolumn{1}{r}{15-22.} & $\isa{ground{\uscore}repl{\uscore}fm}(\phi)$ & one for each formula $\phi$ marked with $\dagger$ \\
  \bottomrule
\end{tabular}
\caption{Replacement Instances used in our mechanization}
\label{tab:instances1}
\end{threeparttable}
\end{table}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "independence_ch_isabelle"
%%% ispell-local-dictionary: "american"
%%% End: