### 2022-10-13 notes/chow.tex
Some words stemming from comments by Tim Chow

Elaboration on comments by Timothy Y. Chow
===========================================

On the relative consistency of $\neg\CH$
----------------------------------------

It is perhaps important to make clear that we have not obtained the
relative consistency of $\neg\CH$, and we are
actually not aiming for that in the short term. In our context, going
for the plain consistency relative to $\ZF$ seems off the mark, since we
can not weaken our base theory (which is essentially equivalent to $\ZF$)%% ,
%% and we catch our tail.
Even if we intended to do so, the standard route to use ctms for a
relative consistency proof (through the Reflection Theorem) is
prohibitive for us, since our metatheory does not have the required
induction principles (on Isabelle/ZF formulas---in contrast to
formulas coded as sets, as in our presentation).

Work by Mathias
---------------

A summary is offered in \cite[Sect.~6]{kanamori-mathias}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In \cite[Sect.~1]{mathias:hal-01188043}, models $M$ of Zermelo Set
Theory are offered that each of the inclusions $M\subseteq M[G]$ and
$M\supseteq M[G]$ fail, where the poset $\pP$ is the trivial
$\{\1 \}$. In one of them, $K$, we have $\om\in K \sm K[G]$, hence the
ordinals do not coincide.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In \cite{mathias-provident}, a reasonably minimal fragment
$\mathsc{Prov}$ of $\ZF$ that allows to do set forcing is identified,
and transitive sets satisfying it are called
\emph{provident}. Existence of rank and of transitive closure are
among them; hence their appearance in our list seems
justified. Nevertheless, $\mathsc{Prov}$ is even a restriction of
Kripke-Platek Set Theory, and thus it does not include neither
Powerset nor full Separation. The detailed theory of provident sets is
developed in \cite{mathias-bowler-gentle}. % Its Section~0 is very
% readable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%