### 2022-10-27 Independence_CH/independence_ch_isabelle.tex
Abstract and capitalization of Zst

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%\usepackage[utf8]{inputenc}
%\usepackage[final]{microtype}
\usepackage{isabelle_indepCH,isabellesym_indepCH}
\usepackage{booktabs,array,threeparttable}
\input{header-indepCH}
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
\hypersetup{
  colorlinks,
  urlcolor={blue},
  linkcolor={blue!50!black},
  citecolor={blue!50!black},
}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\input{unicode_support}
\usepackage{array}
\newcounter{replInstCount}
\setcounter{replInstCount}{0}

\newcounter{LamReplCount}
\setcounter{LamReplCount}{0}
\begin{document}
%
\title{The formal verification of the ctm approach to forcing%Some lessons after the formalization of the ctm approach to forcing%
  \thanks{Supported by Secyt-UNC projects 33620180100855CB and 33620180100465CB, and Conicet.}%
}
%
\titlerunning{Formalization of ctm forcing}%Lessons after formalizing ctm forcing}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Emmanuel Gunther\inst{1} \and
Miguel Pagano\inst{1} \and \\
Pedro Sánchez Terraf\inst{1,2}%\orcidID{0000-0003-3928-6942}
\and
Matías Steinberg\inst{1}
}
%
\authorrunning{E.~Gunther, M.~Pagano, P.~Sánchez Terraf, M.~Steinberg}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universidad Nacional de C\'ordoba. 
  \\  Facultad de Matem\'atica, Astronom\'{\i}a,  F\'{\i}sica y
  Computaci\'on.
  \and
    Centro de Investigaci\'on y Estudios de Matem\'atica (CIEM-FaMAF),
    Conicet. C\'ordoba. Argentina. \\
    \email{\{gunther,sterraf\}@famaf.unc.edu.ar\\
        \{miguel.pagano,matias.steinberg\}@unc.edu.ar}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  We discuss some highlights of our computer-verified
  proof of the construction, given a countable transitive set-model $M$
  of $\ZFC$, of generic extensions  satisfying $\ZFC+\neg\CH$ and $\ZFC+\CH$.
  Moreover, let $\calR$ be the set of instances of the Axiom of
  Replacement. We isolated a 21-element subset $\Omega\sbq\calR$ and
  defined $\calF:\calR\to\calR$
  such that for every $\Phi\sbq\calR$ and $M$-generic $G$,
  $M\models \ZC \cup \calF\dimg\Phi \cup \Omega$ implies
  $M[G]\models \ZC \cup \Phi \cup \{ \neg \CH \}$, where $\ZC$ is
  Zermelo set theory with Choice.
  %% We also obtained the formulas yielded by the Forcing Definability Theorem
  %% explicitly.

  To achieve this, we worked in the proof assistant \emph{Isabelle},
  basing our development on the Isabelle/ZF library by L.~Paulson and
  others.

  \keywords{forcing \and Isabelle/ZF \and countable transitive models
    \and continuum hypothesis
    \and proof assistants \and interactive theorem provers \and generic extension}
\end{abstract}
%
%
%
\input{intro}

\input{proof_assists_isabelle}

\input{non-absolute}

\input{forcing}

\input{sample_proof}

\input{main_results}

\input{related}

\input{lessons}

\input{conclusion}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
%\bibliographystyle{splncs04}
\bibliographystyle{mi-estilo-else}
\bibliography{independence_ch_isabelle}

\appendix

\input{main_defs}

\input{discipline}

\input{cofinality}

\input{axioms_metatheory}

\input{lam_replacement}

\input{replacement_instances}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% ispell-local-dictionary: "american"
%%% End: